When a continuous variable has a right-skewed distribution
(a longer tail to the right than to the left),
a logarithmic transformation often makes the distribution
more symmetric.
A log-transformed variable also has a multiplicative interpretation:
an increase of 1 in $\log(x)$ corresponds to multiplying $x$ by $e \approx 2.72$,
because $\log(x) + 1 = \log(e \cdot x)$.

The WCGS dataset already contains log-transformed versions
of two variables:

- `lnsbp`: $\log(\text{SBP})$;
- `lnwght`: $\log(\text{weight})$.

::: {#fig-hist-sbp-log layout-ncol="2"}

::: {#fig-hist-sbp-raw}

```{r}
wcgs |>
  ggplot2::ggplot() +
  ggplot2::aes(x = sbp) +
  ggplot2::geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  ggplot2::labs(x = "Systolic BP (mmHg)", y = "Count")
```

Raw scale

:::

::: {#fig-hist-sbp-logscale}

```{r}
wcgs |>
  ggplot2::ggplot() +
  ggplot2::aes(x = lnsbp) +
  ggplot2::geom_histogram(bins = 30, fill = "coral", color = "white") +
  ggplot2::labs(x = "log(Systolic BP)", y = "Count")
```

Log scale

:::

Distribution of systolic blood pressure in the WCGS, on two scales

:::

```{r}
#| label: skewness-sbp
#| code-fold: true
# sample skewness: mean cubed deviation, divided by the cubed SD
skewness <- function(x) {
  x <- x[!is.na(x)]
  mean((x - mean(x))^3) / sd(x)^3
}
sbp_skew <- c(
  raw = skewness(wcgs$sbp),
  log = skewness(wcgs$lnsbp)
)
sbp_skew
```

The log-transformed SBP is less skewed than the raw SBP
(sample skewness `r round(sbp_skew[["log"]], 2)`
versus `r round(sbp_skew[["raw"]], 2)`),
but still not symmetric.
Whether to transform a variable in a regression model depends on
the assumptions of that model and on the scientific question.
