::: {#exm-bayes-survival}
#### Exponential survival regression fitted by Bayesian inference

We simulate $N = 200$ right-censored exponential survival times,
with a binary covariate $x_i$ (say, treatment)
and hazard $\haz_i = \expf{\beta_0 + \beta_1 x_i}$,
where the log baseline hazard is $\beta_0 = \log 0.05$
and the log hazard ratio is $\beta_1 = -0.7$.
With event indicator $\delta_i$ (1 for an event, 0 for a censored time)
and follow-up time $t_i$,
the exponential density is $\haz_i e^{-\haz_i t}$
and its survival function is $e^{-\haz_i t}$,
so observation $i$ contributes the log-likelihood

$$
\ell_i \eqdef \delta_i \log \haz_i - \haz_i t_i.
$$

JAGS has no built-in distribution for this contribution,
so we fit it with the zeros trick.

```{r}
#| label: jags-survival
set.seed(2026)
n_obs <- 200L
x <- stats::rbinom(n_obs, size = 1, prob = 0.5)
event_time <- stats::rexp(n_obs, rate = exp(log(0.05) - 0.7 * x))
censor_time <- stats::rexp(n_obs, rate = 0.03)
follow_up <- pmin(event_time, censor_time)
event <- as.integer(event_time <= censor_time)

survival_model <- "
model {
  C <- 10000
  for (i in 1:N) {
    log(lambda[i]) <- beta0 + beta1 * x[i]
    loglik[i] <- event[i] * log(lambda[i]) - lambda[i] * follow_up[i]
    zeros[i] ~ dpois(C - loglik[i])
  }
  beta0 ~ dnorm(0, 0.01)
  beta1 ~ dnorm(0, 0.01)
  HR <- exp(beta1)
}
"
inits <- list(
  list(.RNG.name = "base::Mersenne-Twister", .RNG.seed = 1),
  list(.RNG.name = "base::Mersenne-Twister", .RNG.seed = 2)
)
fit <- rjags::jags.model(
  textConnection(survival_model),
  data = list(
    follow_up = follow_up, x = x, event = event,
    N = n_obs, zeros = rep(0, n_obs)
  ),
  inits = inits,
  n.chains = 2,
  quiet = TRUE
)
stats::update(fit, n.iter = 1000, progress.bar = "none")
draws <- rjags::coda.samples(
  fit,
  variable.names = c("beta0", "beta1", "HR"),
  n.iter = 4000,
  progress.bar = "none"
)
draws_matrix <- as.matrix(draws)
cbind(
  mean = colMeans(draws_matrix),
  t(apply(draws_matrix, 2, stats::quantile, probs = c(0.025, 0.975))),
  Rhat = coda::gelman.diag(draws, multivariate = FALSE)$psrf[, "Point est."]
) |>
  round(3)
```

About `r round(100 * mean(event))`% of the follow-up times end in an event.
The 95% credible interval for $\beta_1$ contains the true $-0.7$,
though the posterior mean is some distance from it.
The maximum likelihood estimates, from the same log-likelihood,
show that the gap is sampling variation in this data set
rather than an effect of the prior:

```{r}
#| label: mle-survival
neg_loglik <- function(beta) {
  lambda <- exp(beta[1] + beta[2] * x)
  -sum(event * log(lambda) - lambda * follow_up)
}
stats::optim(c(0, 0), neg_loglik)$par |>
  stats::setNames(c("beta0", "beta1")) |>
  round(3)
```

The monitored $e^{\beta_1}$ gives the hazard ratio and its credible interval directly.
:::
