::: {#exm-ppp-value}
#### Posterior predictive p-values for the race spread

From the simulations of @exm-posterior-predictive-check:

::: {.panel-tabset group="language"}

##### R

```{r}
#| label: ppp-values
#| code-fold: false
ppp_values <- tapply(
  ppc_draws$t_rep >= ppc_draws$t_obs,
  ppc_draws$data_set,
  mean
)
ppp_values
```

##### Python

```{python}
#| label: ppp-values-py
#| code-fold: false
{name: (t_rep >= t_obs).mean() for name, (t_rep, t_obs) in ppc_draws.items()}
```

:::

```{r}
#| label: ppp-verdicts
#| include: false
ppp_verdict <- function(p) {
  if (p < 0.05 || p > 0.95) {
    "so the model fails it"
  } else {
    "so the model passes this check"
  }
}
```

R and Python draw different random numbers,
so their p-values differ a little;
the text uses the R results.
For the simulated data, $p_B$ is `r round(ppp_values[["simulated"]], 3)`,
`r ppp_verdict(ppp_values[["simulated"]])`.
For the real data, $p_B$ is `r round(ppp_values[["real (birthwt)"]], 3)`,
`r ppp_verdict(ppp_values[["real (birthwt)"]])`.
The next step would be a model that lets the smoking probability depend on race,
such as the [Bayesian logistic regression](bayesian-examples.qmd#exm-bayes-logistic)
fitted on the companion page.
:::
