::: {#exm-bayes-random-effects}
#### A random-intercept model fitted by Bayesian inference

We simulate $J = 8$ groups of 12 observations each,
with group means drawn from $\ndist{\mu, \tau^2}$
($\mu = 5$, $\tau = 1.5$)
and within-group standard deviation $\sigma = 2$.
The priors are a diffuse $\ndist{0, 100^2}$ for $\mu$
and [flat priors](bayesian-inference.qmd#def-flat-prior) on $(0, 100)$
for the standard deviations $\tau$ and $\sigma$.
Being flat on the standard-deviation scale is a choice:
as [a flat prior on the log-odds](bayesian-inference.qmd#exm-flat-prior-reparam) shows for a probability,
it is not flat on another scale, such as the variance.

```{r}
#| label: jags-random-effects
#| code-fold: false
set.seed(2025)
n_groups <- 8L
n_per_group <- 12L
theta_true <- stats::rnorm(n_groups, mean = 5, sd = 1.5)
group <- rep(seq_len(n_groups), each = n_per_group)
y <- stats::rnorm(n_groups * n_per_group, mean = theta_true[group], sd = 2)

random_intercept_model <- "
model {
  for (i in 1:N) {
    y[i] ~ dnorm(theta[group[i]], 1 / sigma^2)
  }
  for (j in 1:J) {
    theta[j] ~ dnorm(mu, 1 / tau^2)
  }
  mu ~ dnorm(0, 1.0E-4)
  sigma ~ dunif(0, 100)
  tau ~ dunif(0, 100)
}
"
inits <- list(
  list(.RNG.name = "base::Mersenne-Twister", .RNG.seed = 1),
  list(.RNG.name = "base::Mersenne-Twister", .RNG.seed = 2)
)
fit <- rjags::jags.model(
  textConnection(random_intercept_model),
  data = list(y = y, group = group, N = length(y), J = n_groups),
  inits = inits,
  n.chains = 2,
  quiet = TRUE
)
stats::update(fit, n.iter = 2000, progress.bar = "none")
draws <- rjags::coda.samples(
  fit,
  variable.names = c("mu", "tau", "sigma", "theta"),
  n.iter = 10000,
  progress.bar = "none"
)
draws_matrix <- as.matrix(draws)
posterior_summary <- cbind(
  mean = colMeans(draws_matrix),
  t(apply(draws_matrix, 2, stats::quantile, probs = c(0.025, 0.975))),
  Rhat = coda::gelman.diag(draws, multivariate = FALSE)$psrf[, "Point est."]
)
round(posterior_summary[c("mu", "tau", "sigma"), ], 3)
```

The 95% credible intervals contain the true $\mu = 5$, $\tau = 1.5$ and $\sigma = 2$.
The interval for $\tau$ is wide:
eight groups say little about how spread out group means are,
and the posterior reports that uncertainty
instead of a single estimate of the variance component.

Each group's posterior mean lies between its sample mean
and the overall sample mean:

```{r}
#| label: random-effects-shrinkage
#| code-fold: false
sample_means <- as.vector(tapply(y, group, mean))
tibble::tibble(
  group_id = seq_len(n_groups),
  sample_mean = sample_means,
  posterior_mean = posterior_summary[
    paste0("theta[", seq_len(n_groups), "]"), "mean"
  ]
) |>
  dplyr::mutate(overall_mean = mean(y)) |>
  round(2)
```

This is the borrowing of strength described in [distributions and hierarchies](bayesian-inference.qmd#sec-bayes-hierarchies):
group 4, with the largest sample mean,
is pulled furthest toward the overall mean.
:::
