::: {#exm-posterior-predictive-check}
#### Does one smoking probability fit every group?

We check the Beta-Bernoulli model with the race-spread test quantity $T_2$
of @exm-test-quantity,
on two data sets:

- **Simulated data.**
  We draw `r n_mothers` Bernoulli observations with the same true $\prob = `r sim_truth`$ for every mother,
  and attach the race labels of the `birthwt` mothers.
  The model is true for these data.
- **Real data.**
  The smoking indicators of the `birthwt` mothers.

For each data set,
we draw $\prob$ from its posterior under the uniform prior,
draw a replicated data set $\vY^{\mathrm{rep}}$ of the same size given that $\prob$,
and compute $T_2(\vY^{\mathrm{rep}})$,
4000 times.

::: {.panel-tabset group="language"}

##### R

```{r}
#| label: posterior-predictive-check
#| code-fold: false
set.seed(4)
check_data <- list(
  simulated = stats::rbinom(n_mothers, size = 1, prob = sim_truth),
  `real (birthwt)` = smoke_y
)
ppc_draws <- do.call(
  rbind,
  lapply(names(check_data), \(data_name) {
    y <- check_data[[data_name]]
    ab <- beta_posterior(y)
    prob_draws <- stats::rbeta(4000, ab[["a"]], ab[["b"]])
    data.frame(
      data_set = data_name,
      t_rep = vapply(
        prob_draws,
        \(p) t_spread(stats::rbinom(n_mothers, size = 1, prob = p)),
        numeric(1)
      ),
      t_obs = t_spread(y)
    )
  })
)
ppc_draws$data_set <- factor(ppc_draws$data_set, levels = names(check_data))
ppc_obs <- unique(ppc_draws[c("data_set", "t_obs")])
ppc_obs
```

##### Python

```{python}
#| label: posterior-predictive-check-py
#| code-fold: false
import numpy as np
import pandas as pd

birthwt = pd.read_csv("data/birthwt.csv")
race = birthwt["race"].to_numpy()
smoke_y = birthwt["smoke"].to_numpy()
n_mothers = len(smoke_y)
sim_truth = 0.3


def t_spread(y):
    group_means = pd.Series(y).groupby(race).mean()
    return group_means.max() - group_means.min()


rng = np.random.default_rng(4)
check_data = {
    "simulated": rng.binomial(1, sim_truth, size=n_mothers),
    "real (birthwt)": smoke_y,
}
ppc_draws = {}
for name, y in check_data.items():
    post_a = 1 + y.sum()
    post_b = 1 + len(y) - y.sum()
    prob_draws = rng.beta(post_a, post_b, size=4000)
    t_rep = np.array([t_spread(rng.binomial(1, p, size=n_mothers))
                      for p in prob_draws])
    ppc_draws[name] = (t_rep, t_spread(y))

{name: t_obs for name, (t_rep, t_obs) in ppc_draws.items()}
```

:::

{{< slidebreak >}}

@fig-posterior-predictive-check compares the replicated and observed values of $T_2$.

::: {#fig-posterior-predictive-check}

::: {.panel-tabset group="language"}

##### R

```{r}
#| label: posterior-predictive-check-plot
#| fig-alt: !expr paste0(
#|   "Two side-by-side histograms of the spread in smoking proportions ",
#|   "across race groups, in data sets replicated from the posterior. ",
#|   "In the left panel (simulated data), the observed spread of ",
#|   round(ppc_obs$t_obs[1], 2),
#|   ", marked by a dashed vertical line, falls inside the histogram. ",
#|   "In the right panel (real birthwt data), the observed spread of ",
#|   round(ppc_obs$t_obs[2], 2),
#|   " lies to the right of nearly all the replicated values."
#|   )
ggplot2::ggplot(ppc_draws, ggplot2::aes(t_rep)) +
  ggplot2::geom_histogram(bins = 40) +
  ggplot2::geom_vline(
    data = ppc_obs,
    ggplot2::aes(xintercept = t_obs),
    linetype = "dashed"
  ) +
  ggplot2::facet_wrap(~data_set) +
  ggplot2::labs(
    x = "spread of smoking proportions across race groups",
    y = "number of simulations"
  )
```

##### Python

```{python}
#| label: posterior-predictive-check-plot-py
#| code-summary: "Show Python code"
#| fig-alt: Two side-by-side histograms of the spread in smoking proportions
#|   across race groups, in data sets replicated from the posterior.
#|   In each panel a dashed vertical line marks the observed spread.
#|   In the left panel (simulated data) it falls inside the histogram;
#|   in the right panel (real birthwt data) it lies to the right of
#|   nearly all the replicated values.
import matplotlib.pyplot as plt

fig, axes = plt.subplots(1, 2, figsize=(8, 3.5), sharey=True)
for ax, (name, (t_rep, t_obs)) in zip(axes, ppc_draws.items()):
    ax.hist(t_rep, bins=40, color="grey")
    ax.axvline(t_obs, color="black", linestyle="--")
    ax.set_title(name)
    ax.set_xlabel("spread of smoking proportions across race groups")
axes[0].set_ylabel("number of simulations")
plt.tight_layout()
plt.show()
```

:::

Posterior predictive distributions of $T_2$,
the spread of the smoking proportions across the three race groups,
from 4000 replicated data sets under the Beta-Bernoulli model with a uniform prior.
The dashed lines are the observed values.
Left: `r n_mothers` simulated observations with the same true $\prob = `r sim_truth`$ for every mother.
Right: the `birthwt` data.
:::

{{< slidebreak >}}

For the simulated data,
the observed spread looks like the replicated spreads,
as it should, because the model generated those data.
For the real data,
the observed spread is larger than almost every replicated spread:
the smoking proportion really does differ by race in these data,
and a model with one $\prob$ for everyone cannot reproduce that.
:::
