::: {#exm-normal-credible-vs-confidence}
#### Credible and confidence intervals for a Gaussian mean

We simulate $n = 20$ observations from the model of @exm-normal-normal,
with true mean $\mu = 2$,
and compute both a 95% confidence interval,
$\bar x \pm 1.96 / \sqrt{n}$ (the variance is known to be 1),
and the equal-tailed 95% credible interval
from the $\ndist{\frac{n}{n+1}\bar{x},\, \frac{1}{n+1}}$ posterior:

```{r}
#| label: normal-credible-vs-confidence
set.seed(1)
mu_true <- 2
n <- 20
x <- rnorm(n = n, mean = mu_true, sd = 1)
xbar <- mean(x)
ci_freq <- xbar + stats::qnorm(c(0.025, 0.975)) / sqrt(n)
post_mean <- n / (n + 1) * xbar
post_sd <- sqrt(1 / (n + 1))
ci_bayes <- stats::qnorm(c(0.025, 0.975), mean = post_mean, sd = post_sd)
rbind(confidence = ci_freq, credible = ci_bayes) |>
  round(3) |>
  `colnames<-`(c("lower", "upper"))
```

The sample mean is $\bar x = `r round(xbar, 3)`$.
The credible interval is slightly narrower than the confidence interval,
because the prior adds information,
and shifted toward the prior mean $0$.

@fig-normal-prior-lik-post shows how the posterior combines prior and data.
Completing the square in the likelihood of @exm-normal-normal
as a function of $\mu$ alone
shows that the likelihood is proportional to a $\ndist{\bar x, 1/n}$ density,
which is the curve drawn for it.

::: {#fig-normal-prior-lik-post}

```{r}
#| code-fold: true
#| fig-alt: >-
#|   Three bell-shaped curves over the mean mu: a wide prior centered at 0,
#|   a narrow likelihood centered near 2.2, and a posterior just as narrow,
#|   centered slightly closer to 0 than the likelihood.
mu_grid <- seq(-3, 5, length.out = 801)
curves <- tibble::tibble(
  mu = rep(mu_grid, times = 3),
  curve = rep(c("prior", "likelihood", "posterior"), each = length(mu_grid)),
  density = c(
    stats::dnorm(mu_grid, mean = 0, sd = 1),
    stats::dnorm(mu_grid, mean = xbar, sd = 1 / sqrt(n)),
    stats::dnorm(mu_grid, mean = post_mean, sd = post_sd)
  )
)
ggplot2::ggplot(curves) +
  ggplot2::aes(x = mu, y = density, colour = curve) +
  ggplot2::geom_line() +
  ggplot2::labs(x = expression(mu), y = "density", colour = NULL)
```

The $\ndist{0, 1}$ prior, the likelihood (scaled to integrate to 1),
and the posterior for the mean $\mu$,
for the data of @exm-normal-credible-vs-confidence.
:::

Because $\mu$ has a distribution,
the Bayesian can also compute the posterior probability
that $\mu$ lies in the realized *confidence* interval:

```{r}
#| label: normal-posterior-prob-ci
pr_in_ci <- stats::pnorm(ci_freq, mean = post_mean, sd = post_sd) |> diff()
round(pr_in_ci, 3)
```

That probability is `r round(pr_in_ci, 3)`, not 0.95:
under this prior,
the confidence interval sits slightly too far from $0$.
A frequentist cannot assign any probability to the event
that $\mu$ lies in one realized interval,
since both are fixed.
:::
