::: {#exm-bayes-logistic}
#### Bayesian logistic regression

We simulate $N = 200$ observations from a logistic model
with one continuous predictor,
intercept $\beta_0 = -0.5$ and slope $\beta_1 = 1.2$,
and place a diffuse $\ndist{0, 10^2}$ prior on each coefficient
(precision $0.01$ in JAGS).
The model also monitors the odds ratio $e^{\beta_1}$
for a one-unit increase in $x$.

```{r}
#| label: jags-logistic
set.seed(2024)
n_obs <- 200L
x <- stats::rnorm(n_obs)
y <- stats::rbinom(n_obs, size = 1, prob = stats::plogis(-0.5 + 1.2 * x))

logistic_model <- "
model {
  for (i in 1:N) {
    y[i] ~ dbern(pi[i])
    logit(pi[i]) <- beta0 + beta1 * x[i]
  }
  beta0 ~ dnorm(0, 0.01)
  beta1 ~ dnorm(0, 0.01)
  OR <- exp(beta1)
}
"
inits <- list(
  list(.RNG.name = "base::Mersenne-Twister", .RNG.seed = 1),
  list(.RNG.name = "base::Mersenne-Twister", .RNG.seed = 2)
)
fit <- rjags::jags.model(
  textConnection(logistic_model),
  data = list(y = y, x = x, N = n_obs),
  inits = inits,
  n.chains = 2,
  quiet = TRUE
)
stats::update(fit, n.iter = 1000, progress.bar = "none")
draws <- rjags::coda.samples(
  fit,
  variable.names = c("beta0", "beta1", "OR"),
  n.iter = 4000,
  progress.bar = "none"
)
draws_matrix <- as.matrix(draws)
cbind(
  mean = colMeans(draws_matrix),
  t(apply(draws_matrix, 2, stats::quantile, probs = c(0.025, 0.975))),
  Rhat = coda::gelman.diag(draws, multivariate = FALSE)$psrf[, "Point est."]
) |>
  round(3)
```

Each 95% credible interval contains its coefficient's true value,
and every $\hR$ is near 1.
As [comparing MCMC estimates to maximum likelihood](mcmc.qmd#sec-mcmc-vs-mle) leads us to expect,
the posterior means are close to the maximum likelihood estimates:

```{r}
#| label: glm-logistic
stats::glm(y ~ x, family = stats::binomial()) |>
  stats::coef() |>
  round(3)
```
:::
