::: {#exm-bayes-bma}
#### A BIC approximation to Bayesian model averaging

Marginal likelihoods are hard to compute,
but with equal prior probabilities for the models,
the [Bayesian information criterion](https://morrison-lab.github.io/rme/chapters/Linear-models-overview.html#def-bic)
gives a large-sample approximation to the posterior model probabilities
[@schwarz1978estimating]:

$$
\p(M_k \mid \vy) \approx
\frac{\expf{-\tfrac{1}{2}\operatorname{BIC}_k}}{\sumto{l}{K} \expf{-\tfrac{1}{2}\operatorname{BIC}_l}}.
$$

We simulate data in which only $x_1$ and $x_2$ affect the outcome,
fit a linear regression for every subset of the three predictors $x_1$, $x_2$ and $x_3$,
and compute these approximate posterior model probabilities:

```{r}
#| label: bma-models
#| code-fold: false
set.seed(2027)
n_obs <- 120L
dat <- tibble::tibble(
  x1 = stats::rnorm(n_obs),
  x2 = stats::rnorm(n_obs),
  x3 = stats::rnorm(n_obs)
) |>
  dplyr::mutate(y = 1 + 0.8 * x1 + 0.5 * x2 + stats::rnorm(n_obs))

predictors <- c("x1", "x2", "x3")
subsets <- lapply(0:3, function(k) {
  utils::combn(predictors, k, simplify = FALSE)
}) |>
  unlist(recursive = FALSE)
models <- lapply(subsets, function(vars) {
  rhs <- if (length(vars) > 0) paste(vars, collapse = " + ") else "1"
  stats::lm(stats::as.formula(paste("y ~", rhs)), data = dat)
})
bic <- vapply(models, stats::BIC, numeric(1))
weight <- exp(-0.5 * (bic - min(bic)))
weight <- weight / sum(weight)

tibble::tibble(
  model = vapply(
    subsets,
    function(s) if (length(s) > 0) paste(s, collapse = " + ") else "(none)",
    character(1)
  ),
  weight = round(weight, 3)
)
```

Subtracting the smallest BIC before exponentiating
leaves the normalized weights unchanged
and avoids numerical underflow.
The [posterior inclusion probability](#def-pip) of each predictor
sums the weights of the models that contain it:

```{r}
#| label: bma-pip
#| code-fold: false
includes <- function(v) vapply(subsets, function(s) v %in% s, logical(1))
vapply(predictors, function(v) sum(weight[includes(v)]), numeric(1)) |>
  round(3)
```

For the model-averaged slope of $x_1$,
we approximate its posterior mean within each model by the least-squares estimate,
taken as $0$ in models that exclude $x_1$,
and average with the posterior model probabilities as weights:

```{r}
#| label: bma-slope
#| code-fold: false
beta_x1 <- vapply(models, function(m) {
  cf <- stats::coef(m)
  if ("x1" %in% names(cf)) cf[["x1"]] else 0
}, numeric(1))
c(
  bma_slope_x1 = sum(weight * beta_x1),
  least_squares_x1_x2 = stats::coef(stats::lm(y ~ x1 + x2, data = dat))[["x1"]]
) |>
  round(3)
```

The predictors that affect the outcome, $x_1$ and $x_2$,
have inclusion probabilities near 1,
and the irrelevant $x_3$ a much smaller one.
The models without $x_1$ get almost no weight,
so the model-averaged slope of $x_1$ is close to its least-squares estimate
in the model with $x_1$ and $x_2$ only.
Both differ from the true slope $0.8$ by sampling variation in this data set,
not because of the averaging.
:::
