::: {#exm-prior-posterior-figure}
#### Prior and posterior, with and without a known truth

Plotting the prior and posterior densities together
shows how much the data moved our beliefs.
We use the Beta-Bernoulli model of @exm-beta-bernoulli
with the uniform $\dBetaf{1, 1}$ prior of @exm-prior
on two data sets:

- **Simulated data.**
  We choose the true value of $\prob$ ourselves,
  draw Bernoulli observations with that probability,
  and then check where the posterior puts the truth.
- **Real data.**
  We use the `birthwt` data on `r nrow(MASS::birthwt)` births
  at Baystate Medical Center in Springfield, Massachusetts, in 1986
  [@hosmer1989applied; @venables2002modern, Section 7.2].
  Here $\prob$ is the probability that a mother in this population smoked during pregnancy,
  and its true value is unknown.
  The data ship with R's `MASS` package as `MASS::birthwt`;
  the Python code reads [`data/birthwt.csv`](data/birthwt.csv), a copy of the same table on this site,
  which is also posted as a CSV file by the
  [Rdatasets project](https://vincentarelbundock.github.io/Rdatasets/csv/MASS/birthwt.csv).

::: {.panel-tabset group="language"}

##### R

```{r}
#| label: prior-posterior-setup
#| code-fold: false
prior_a <- 1
prior_b <- 1

set.seed(1)
sim_truth <- 0.3
sim_y <- stats::rbinom(50, size = 1, prob = sim_truth)
smoke_y <- MASS::birthwt$smoke

beta_posterior <- function(y, a = prior_a, b = prior_b) {
  c(a = a + sum(y), b = b + length(y) - sum(y))
}

post_params <- rbind(
  simulated = beta_posterior(sim_y),
  `real (birthwt)` = beta_posterior(smoke_y)
)
post_summary <- data.frame(
  n = c(length(sim_y), length(smoke_y)),
  successes = c(sum(sim_y), sum(smoke_y)),
  post_a = post_params[, "a"],
  post_b = post_params[, "b"],
  mean = post_params[, "a"] / rowSums(post_params),
  lower = stats::qbeta(0.025, post_params[, "a"], post_params[, "b"]),
  upper = stats::qbeta(0.975, post_params[, "a"], post_params[, "b"])
)
post_summary
```

##### Python

```{python}
#| label: prior-posterior-setup-py
#| code-fold: false
import numpy as np
import pandas as pd
from scipy import stats

prior_a, prior_b = 1, 1

rng = np.random.default_rng(1)
sim_truth = 0.3
sim_y = rng.binomial(1, sim_truth, size=50)
smoke_y = pd.read_csv("data/birthwt.csv")["smoke"].to_numpy()


def beta_posterior(y, a=prior_a, b=prior_b):
    return a + y.sum(), b + len(y) - y.sum()


rows = []
for name, y in {"simulated": sim_y, "real (birthwt)": smoke_y}.items():
    post_a, post_b = beta_posterior(y)
    rows.append({
        "data": name,
        "n": len(y),
        "successes": y.sum(),
        "post_a": post_a,
        "post_b": post_b,
        "mean": post_a / (post_a + post_b),
        "lower": stats.beta.ppf(0.025, post_a, post_b),
        "upper": stats.beta.ppf(0.975, post_a, post_b),
    })
pd.DataFrame(rows).set_index("data")
```

:::

R and Python use different random-number generators,
so their simulated draws differ, and their summaries can differ too;
the numbers in the text below come from the R code.

{{< slidebreak >}}

@fig-prior-posterior plots each prior and posterior density.

::: {#fig-prior-posterior}

::: {.panel-tabset group="language"}

##### R

```{r}
#| label: prior-posterior-plot
#| fig-alt: !expr paste0(
#|   "Two side-by-side panels of densities for the probability of a success, ",
#|   "each on the interval from 0 to 1. ",
#|   "In both panels the uniform prior is a flat line at height 1, ",
#|   "and the posterior is a hump. ",
#|   "In the left panel (simulated data, ", length(sim_y), " observations), ",
#|   "a dashed vertical line at the true value ", sim_truth,
#|   " passes through the posterior hump, which is centred near ",
#|   round(post_summary["simulated", "mean"], 2), ". ",
#|   "In the right panel (real birthwt data, ", length(smoke_y), " births), ",
#|   "the posterior hump is narrower, centred near ",
#|   round(post_summary["real (birthwt)", "mean"], 2),
#|   ", and no true value is marked."
#|   )
prob_grid <- seq(0.001, 0.999, length.out = 500)
density_curves <- dplyr::bind_rows(
  lapply(rownames(post_params), \(data_set) {
    dplyr::bind_rows(
      tibble::tibble(
        data_set = data_set,
        prob = prob_grid,
        distribution = "prior",
        density = stats::dbeta(prob_grid, prior_a, prior_b)
      ),
      tibble::tibble(
        data_set = data_set,
        prob = prob_grid,
        distribution = "posterior",
        density = stats::dbeta(
          prob_grid,
          post_params[data_set, "a"],
          post_params[data_set, "b"]
        )
      )
    )
  })
)
density_curves$data_set <- factor(
  density_curves$data_set,
  levels = rownames(post_params)
)
density_curves$distribution <- factor(
  density_curves$distribution,
  levels = c("prior", "posterior")
)

ggplot2::ggplot(
  density_curves,
  ggplot2::aes(prob, density, colour = distribution, linetype = distribution)
) +
  ggplot2::geom_line(linewidth = 0.8) +
  ggplot2::geom_vline(
    data = tibble::tibble(
      data_set = factor("simulated", levels = rownames(post_params)),
      truth = sim_truth
    ),
    ggplot2::aes(xintercept = truth),
    linetype = "dashed"
  ) +
  ggplot2::facet_wrap(~data_set) +
  ggplot2::labs(
    x = "probability of a success",
    y = "density",
    colour = NULL,
    linetype = NULL
  )
```

##### Python

```{python}
#| label: prior-posterior-plot-py
#| code-summary: "Show Python code"
#| fig-alt: Two side-by-side panels of densities for the probability of a success,
#|   each on the interval from 0 to 1.
#|   In both panels the uniform prior is a flat line,
#|   and the posterior is a hump.
#|   In the left panel (simulated data)
#|   a dotted vertical line marks the true value.
#|   In the right panel (real birthwt data)
#|   the posterior hump is narrower, and no true value is marked.
import matplotlib.pyplot as plt
import numpy as np
from scipy import stats

prob_grid = np.linspace(0.001, 0.999, 500)
data_sets = {"simulated": sim_y, "real (birthwt)": smoke_y}
fig, axes = plt.subplots(1, 2, figsize=(8, 3.5), sharey=True)
for ax, (name, y) in zip(axes, data_sets.items()):
    post_a, post_b = beta_posterior(y)
    ax.plot(prob_grid, stats.beta.pdf(prob_grid, prior_a, prior_b),
            label="prior")
    ax.plot(prob_grid, stats.beta.pdf(prob_grid, post_a, post_b),
            linestyle="--", label="posterior")
    if name == "simulated":
        ax.axvline(sim_truth, color="black", linestyle=":", label="truth")
    ax.set_title(name)
    ax.set_xlabel("probability of a success")
axes[0].set_ylabel("density")
axes[0].legend()
plt.tight_layout()
plt.show()
```

:::

Prior and posterior densities of $\prob$
under a uniform $\dBetaf{1, 1}$ prior.
Left: `r post_summary["simulated", "n"]` simulated Bernoulli observations,
drawn with true $\prob = `r sim_truth`$ (vertical line).
Right: whether each of the `r post_summary["real (birthwt)", "n"]` mothers
in the `birthwt` data smoked during pregnancy;
the true $\prob$ is unknown.
:::

{{< slidebreak >}}

In the simulated panel,
`r post_summary["simulated", "successes"]` of the
`r post_summary["simulated", "n"]` draws were successes,
so the posterior is
$\dBetaf{`r post_summary["simulated", "post_a"]`, `r post_summary["simulated", "post_b"]`}$.
Its mean is `r round(post_summary["simulated", "mean"], 3)`,
and its equal-tailed 95% [credible interval](#def-credible-interval)
runs from `r round(post_summary["simulated", "lower"], 3)`
to `r round(post_summary["simulated", "upper"], 3)`,
which `r if (post_summary["simulated", "lower"] <= sim_truth && sim_truth <= post_summary["simulated", "upper"]) "contains" else "misses"`
the true value `r sim_truth`.
Because we chose the truth,
we can check the posterior against it.

In the real panel,
`r post_summary["real (birthwt)", "successes"]` of the
`r post_summary["real (birthwt)", "n"]` mothers smoked,
so the posterior is
$\dBetaf{`r post_summary["real (birthwt)", "post_a"]`, `r post_summary["real (birthwt)", "post_b"]`}$,
with mean `r round(post_summary["real (birthwt)", "mean"], 3)`
and 95% credible interval
from `r round(post_summary["real (birthwt)", "lower"], 3)`
to `r round(post_summary["real (birthwt)", "upper"], 3)`.
No true value is available to check it against:
the posterior is all we have to describe what the data say about $\prob$.

In both panels,
the posterior is much narrower than the flat prior,
and the real-data posterior,
based on more observations,
is the narrower of the two.
:::
