::: {#exm-prior-predictive-check}
#### Checking two priors for the smoking probability

We compare two priors for $\prob$ in the `birthwt` data:

- the uniform $\dBetaf{1, 1}$ prior of @exm-prior;
- a $\dBetaf{1, 19}$ prior,
  with prior mean $1 / (1 + 19) = `r 1 / (1 + 19)`$,
  which an analyst might choose if they believed that smoking during pregnancy was rare.

For each prior,
we draw $\prob$ from the prior,
then draw a count of smokers $T_1(\vY)$ among `r n_mothers` mothers given that $\prob$.

::: {.panel-tabset group="language"}

##### R

```{r}
#| label: prior-predictive-check
#| code-fold: false
check_priors <- list(
  `uniform Beta(1, 1)` = c(a = 1, b = 1),
  `Beta(1, 19)` = c(a = 1, b = 19)
)
set.seed(3)
prior_pred_counts <- dplyr::bind_rows(
  lapply(names(check_priors), \(prior_name) {
    ab <- check_priors[[prior_name]]
    prob_draws <- stats::rbeta(4000, ab[["a"]], ab[["b"]])
    tibble::tibble(
      prior = prior_name,
      count = stats::rbinom(4000, size = n_mothers, prob = prob_draws)
    )
  })
)
prior_pred_counts$prior <- factor(
  prior_pred_counts$prior,
  levels = names(check_priors)
)
prior_tail <- tapply(
  prior_pred_counts$count >= t_count(smoke_y),
  prior_pred_counts$prior,
  mean
)
prior_tail
```

##### Python

```{python}
#| label: prior-predictive-check-py
#| code-fold: false
import numpy as np
import pandas as pd

smoke_y = pd.read_csv("data/birthwt.csv")["smoke"].to_numpy()
n_mothers = len(smoke_y)
check_priors = {"uniform Beta(1, 1)": (1, 1), "Beta(1, 19)": (1, 19)}

rng = np.random.default_rng(3)
prior_pred_counts = {}
for name, (a, b) in check_priors.items():
    prob_draws = rng.beta(a, b, size=4000)
    prior_pred_counts[name] = rng.binomial(n_mothers, prob_draws)

{name: (counts >= smoke_y.sum()).mean()
 for name, counts in prior_pred_counts.items()}
```

:::

The output is the proportion of simulated counts at least as large as the observed count,
`r t_count(smoke_y)`.

{{< slidebreak >}}

@fig-prior-predictive-check plots the simulated counts.

::: {#fig-prior-predictive-check}

::: {.panel-tabset group="language"}

##### R

```{r}
#| label: prior-predictive-check-plot
#| fig-alt: !expr paste0(
#|   "Two side-by-side histograms of the number of smokers among ",
#|   n_mothers, " mothers, simulated from two priors. ",
#|   "Under the uniform Beta(1, 1) prior, the counts spread evenly from 0 to ",
#|   n_mothers, ", and the observed count of ", t_count(smoke_y),
#|   ", marked by a dashed vertical line, sits in the middle. ",
#|   "Under the Beta(1, 19) prior, most counts are small, ",
#|   "and the observed count is far to the right of almost all of them."
#|   )
ggplot2::ggplot(prior_pred_counts, ggplot2::aes(count)) +
  ggplot2::geom_histogram(binwidth = 5, boundary = 0) +
  ggplot2::geom_vline(xintercept = t_count(smoke_y), linetype = "dashed") +
  ggplot2::facet_wrap(~prior) +
  ggplot2::labs(x = "number of smokers", y = "number of simulations")
```

##### Python

```{python}
#| label: prior-predictive-check-plot-py
#| code-summary: "Show Python code"
#| fig-alt: Two side-by-side histograms of the number of smokers,
#|   simulated from two priors.
#|   Under the uniform Beta(1, 1) prior, the counts spread evenly,
#|   and the observed count, marked by a dashed vertical line, sits in the middle.
#|   Under the Beta(1, 19) prior, most counts are small,
#|   and the observed count is far to the right of almost all of them.
import matplotlib.pyplot as plt
import numpy as np

fig, axes = plt.subplots(1, 2, figsize=(8, 3.5), sharey=True)
for ax, (name, counts) in zip(axes, prior_pred_counts.items()):
    ax.hist(counts, bins=np.arange(0, n_mothers + 6, 5), color="grey")
    ax.axvline(smoke_y.sum(), color="black", linestyle="--")
    ax.set_title(name)
    ax.set_xlabel("number of smokers")
axes[0].set_ylabel("number of simulations")
plt.tight_layout()
plt.show()
```

:::

Prior predictive distributions of the number of smokers
among the `r n_mothers` mothers in the `birthwt` data,
from 4000 simulations under each prior.
The dashed line is the observed count, `r t_count(smoke_y)`.
:::

{{< slidebreak >}}

Under the uniform prior,
`r round(prior_tail[["uniform Beta(1, 1)"]], 3)` of the simulated counts
are at least as large as the observed count,
so the data are unsurprising under that prior.
Under the $\dBetaf{1, 19}$ prior,
`r sum(prior_pred_counts$prior == "Beta(1, 19)" & prior_pred_counts$count >= t_count(smoke_y))`
of the 4000 simulated counts are:
that prior and the data disagree,
and the analyst should rethink the prior
before trusting a posterior built on it.
:::
