::: {#exm-hers-ttest}
#### Baseline fasting glucose by treatment group in HERS

We test $H_0: \mu_\text{HT} = \mu_\text{placebo}$
against $H_1: \mu_\text{HT} \neq \mu_\text{placebo}$,
where $\mu_\text{HT}$ and $\mu_\text{placebo}$
are the mean baseline fasting glucose levels
in the hormone therapy and placebo populations.
@fig-hers-boxplot compares the two groups' distributions.

::: {#fig-hers-boxplot}

```{r}
#| code-fold: true
hers |>
  ggplot2::ggplot() +
  ggplot2::aes(x = HT, y = glucose) +
  ggplot2::geom_boxplot() +
  ggplot2::labs(
    x = "Treatment group",
    y = "Fasting glucose (mg/dL)"
  )
```

Baseline fasting glucose by treatment group in HERS

:::

The statistic and degrees of freedom of @def-two-sample-t-test:

```{r}
#| code-fold: false
glucose_ht <- hers |>
  dplyr::filter(HT == "hormone therapy") |>
  dplyr::pull(glucose)
glucose_placebo <- hers |>
  dplyr::filter(HT == "placebo") |>
  dplyr::pull(glucose)
v_ht <- var(glucose_ht) / length(glucose_ht)
v_placebo <- var(glucose_placebo) / length(glucose_placebo)
t_welch <- (mean(glucose_ht) - mean(glucose_placebo)) / sqrt(v_ht + v_placebo)
df_welch <- (v_ht + v_placebo)^2 /
  (v_ht^2 / (length(glucose_ht) - 1) +
     v_placebo^2 / (length(glucose_placebo) - 1))
c(t = t_welch, df = df_welch, p_value = 2 * pt(-abs(t_welch), df_welch))
```

`t.test()` reports the same values:

```{r}
#| code-fold: false
t.test(glucose_ht, glucose_placebo)
```

The p-value is large,
so the data give no evidence against $H_0$.
That result is expected here:
treatment was randomized,
so the groups' baseline means differ only by chance.
:::
