::: {#exm-hers-anova}
#### Fasting glucose by race/ethnicity in HERS

The group sizes, means, and standard deviations
of baseline fasting glucose:

```{r}
#| code-fold: false
hers |>
  dplyr::summarize(
    .by = raceth,
    n = dplyr::n(),
    mean = mean(glucose),
    sd = sd(glucose)
  )
```

The F statistic of @def-one-way-anova, computed step by step:

```{r}
#| code-fold: false
anova_parts <- hers |>
  dplyr::mutate(grand_mean = mean(glucose)) |>
  dplyr::mutate(.by = raceth, group_mean = mean(glucose)) |>
  dplyr::summarize(
    ss_between = sum((group_mean - grand_mean)^2),
    ss_within = sum((glucose - group_mean)^2),
    k = dplyr::n_distinct(raceth),
    n = dplyr::n()
  ) |>
  dplyr::mutate(
    f = (ss_between / (k - 1)) / (ss_within / (n - k)),
    p_value = pf(f, k - 1, n - k, lower.tail = FALSE)
  )
anova_parts
```

$\text{SS}_\text{between}$ is summed here over observations rather than groups:
each observation in group $j$ contributes $(\bar{y}_j - \bar{y})^2$,
which gives the $n_j$ weights of @def-one-way-anova.
`aov()` reports the same sums of squares, F statistic, and p-value:

```{r}
#| code-fold: false
aov(glucose ~ raceth, data = hers) |> summary()
```

The p-value is far below 0.05:
mean fasting glucose differs among the three race/ethnicity groups.
:::

{{< slidebreak >}}

::: {#rem-hers-anova-variances}
#### Unequal variances in the HERS glucose groups

The group standard deviations differ,
from `r round(min(tapply(hers$glucose, hers$raceth, sd)))`
to `r round(max(tapply(hers$glucose, hers$raceth, sd)))` mg/dL,
so the equal-variance condition of @thm-anova-null is questionable.
`oneway.test()` performs Welch's version of the F-test,
which does not assume equal variances,
and reaches the same conclusion:

```{r}
#| code-fold: false
oneway.test(glucose ~ raceth, data = hers)
```
:::
