::: {#exm-hers-slr}
#### Regression of fasting glucose on BMI in HERS

The OLS estimates from @thm-ols-slr,
and the slope from @cor-ols-slope-r,
for the participants with a BMI measurement:

```{r}
#| code-fold: false
hers_bmi <- hers |> dplyr::filter(!is.na(BMI))
x <- hers_bmi$BMI
y <- hers_bmi$glucose
b1 <- sum((x - mean(x)) * (y - mean(y))) / sum((x - mean(x))^2)
c(
  intercept = mean(y) - b1 * mean(x),
  slope = b1,
  slope_from_r = cor(x, y) * sd(y) / sd(x)
)
```

`lm()` gives the same estimates:

```{r}
#| code-fold: false
slr_fit <- lm(glucose ~ BMI, data = hers)
parameters::parameters(slr_fit) |>
  parameters::print_md(digits = 3)
performance::model_performance(slr_fit) |>
  insight::print_md(digits = 3)
```

@fig-hers-slr shows the data with the fitted line.

::: {#fig-hers-slr}

```{r}
ggplot2::ggplot(hers_bmi, ggplot2::aes(x = BMI, y = glucose)) +
  ggplot2::geom_point(alpha = 0.2) +
  ggplot2::geom_abline(
    intercept = coef(slr_fit)[[1]],
    slope = coef(slr_fit)[["BMI"]],
    color = "firebrick"
  ) +
  ggplot2::labs(x = "BMI (kg/m^2)", y = "Fasting glucose (mg/dL)")
```

Fasting glucose against BMI in HERS, with the OLS line

:::

The estimated slope is
$\hcoef{\text{BMI}} = `r round(coef(slr_fit)[["BMI"]], 2)`$ mg/dL per kg/m^2^:
mean fasting glucose is about `r round(coef(slr_fit)[["BMI"]], 1)` mg/dL higher
among participants whose BMI is 1 kg/m^2^ higher.
The t statistic for the slope equals the correlation test statistic
of @exm-hers-cor.
:::
