::: {#exm-hers-cor}
#### Correlation between BMI and fasting glucose in HERS

@fig-hers-scatter plots baseline fasting glucose against BMI.

::: {#fig-hers-scatter}

```{r}
#| code-fold: true
hers |>
  dplyr::filter(!is.na(BMI)) |>
  ggplot2::ggplot() +
  ggplot2::aes(x = BMI, y = glucose) +
  ggplot2::geom_point(alpha = 0.3) +
  ggplot2::geom_smooth(method = "lm", formula = y ~ x) +
  ggplot2::labs(
    x = "BMI (kg/m^2)",
    y = "Fasting glucose (mg/dL)"
  )
```

Baseline fasting glucose against BMI in HERS,
with the least-squares line

:::

The statistic of @def-pearson-test,
using the `r sum(!is.na(hers$BMI))` participants with a BMI measurement:

```{r}
hers_bmi <- hers |> dplyr::filter(!is.na(BMI))
r <- cor(hers_bmi$BMI, hers_bmi$glucose)
n <- nrow(hers_bmi)
t_cor <- r * sqrt(n - 2) / sqrt(1 - r^2)
c(r = r, t = t_cor, p_value = 2 * pt(-abs(t_cor), df = n - 2))
```

`cor.test()` reports the same values,
and a confidence interval for $\rho$:

```{r}
cor.test(hers$BMI, hers$glucose, method = "pearson")
```

The correlation is positive but modest:
glucose tends to be higher at higher BMI,
with wide scatter around the line in @fig-hers-scatter.
:::
