::: {#exm-hers-chisq}
#### Chi-square test of exercise by treatment group in HERS

The expected counts and statistic of @def-chi-square-test
for @tbl-hers-crosstab:

```{r}
#| code-fold: false
observed <- table(hers$exercise, hers$HT)
expected <- outer(rowSums(observed), colSums(observed)) / sum(observed)
x_squared <- sum((observed - expected)^2 / expected)
expected
c(
  x_squared = x_squared,
  p_value = pchisq(x_squared, df = 1, lower.tail = FALSE)
)
```

Every expected count is large,
so the $\chi^2_1$ approximation of @thm-chi-square-null is reasonable.
`chisq.test()` with `correct = FALSE` reports the same values:

```{r}
#| code-fold: false
chisq.test(hers$exercise, hers$HT, correct = FALSE)
```

The p-value is large:
the data give no evidence that exercise depends on treatment group,
as randomization would lead us to expect.
:::

::: notes
For a $2 \times 2$ table, `chisq.test()` applies
Yates' continuity correction by default,
which subtracts 0.5 from each $\abs{O_{ij} - E_{ij}}$ before squaring,
and so gives a smaller statistic than @def-chi-square-test.
`correct = FALSE` turns the correction off.
:::
