::: {#exm-bootstrap-distribution-toy}
#### Bootstrap distribution of a mean of ten glucose values

Take the baseline fasting glucose values
of the first ten HERS participants,
and compute the sample mean of each of $B = 1{,}000$ bootstrap samples:

```{r}
#| code-fold: false
glucose10 <- hers$glucose[1:10]
set.seed(42)
boot_means10 <- replicate(1000, mean(sample(glucose10, replace = TRUE)))
c(
  observed_mean = mean(glucose10),
  bootstrap_se = sd(boot_means10),
  formula_se = sd(glucose10) / sqrt(10),
  divide_by_n_se = sqrt(mean((glucose10 - mean(glucose10))^2) / 10)
)
```

The bootstrap standard error (@def-bootstrap-se)
is close to the usual estimate $s / \sqrt{n}$.
Each bootstrap draw has variance
$\hat\sigma^2 \eqdef \frac{1}{n} \sum_{i=1}^n (x_i - \bar{x})^2$
around $\bar{x}$, and the $n$ draws are independent,
so as $B$ grows the bootstrap standard error of the mean
approaches $\hat\sigma / \sqrt{n}$ (`divide_by_n_se`),
which is smaller than $s / \sqrt{n}$ by the factor $\sqrt{(n-1)/n}$.
@fig-boot-dist-toy shows the bootstrap distribution.

::: {#fig-boot-dist-toy}

```{r}
#| code-fold: true
tibble::tibble(mean = boot_means10) |>
  ggplot2::ggplot() +
  ggplot2::aes(x = mean) +
  ggplot2::geom_histogram(bins = 30, color = "white") +
  ggplot2::geom_vline(xintercept = mean(glucose10), color = "red") +
  ggplot2::labs(x = "Bootstrap mean (mg/dL)", y = "Count")
```

Bootstrap distribution of the mean fasting glucose
of the first ten HERS participants ($B = 1{,}000$).
The red line marks the observed mean.

:::

:::
