::: {#exm-hers-paired}
#### Change in fasting glucose over the first year of HERS

We test whether mean fasting glucose changed
between baseline (`glucose`) and the year-1 visit (`glucose1`),
across both treatment groups.
Participants missing either measurement are dropped.
The t statistic of @def-one-sample-t-test, computed on the differences:

```{r}
#| code-fold: false
glucose_change <- hers |>
  dplyr::filter(!is.na(glucose), !is.na(glucose1)) |>
  dplyr::mutate(d = glucose1 - glucose) |>
  dplyr::pull(d)
n_pairs <- length(glucose_change)
t_paired <- mean(glucose_change) / (sd(glucose_change) / sqrt(n_pairs))
c(
  n = n_pairs,
  mean_change = mean(glucose_change),
  t = t_paired,
  p_value = 2 * pt(-abs(t_paired), df = n_pairs - 1)
)
```

`t.test()` with `paired = TRUE` gives the same statistic,
along with a confidence interval for $\mean_d$:

```{r}
#| code-fold: false
t.test(hers$glucose1, hers$glucose, paired = TRUE)
```

Mean fasting glucose rose by about
`r round(mean(glucose_change), 1)` mg/dL over the year,
and the p-value is far below 0.05.
:::
