::: {#exm-hers-bootstrap}
#### Bootstrap confidence intervals for the slope of SBP on age in HERS

We regress systolic blood pressure (`SBP`) on `age`
by [ordinary least squares](correlation-regression.qmd#def-ols),
and bootstrap the slope,
resampling participants [adapted from @vittinghoff2e, Chapter 3].
The statistic function takes the data and the row indices
of one bootstrap sample:

```{r}
slope_sbp_age <- function(data, indices) {
  fit <- lm(SBP ~ age, data = data[indices, ])
  coef(fit)[["age"]]
}

set.seed(42)
boot_result <- boot::boot(data = hers, statistic = slope_sbp_age, R = 1000)
boot_result
```

The bootstrap standard error is close to
the model-based standard error from `lm()`,
and the three bootstrap intervals are close to
the model-based 95% confidence interval:

```{r}
fit_sbp_age <- lm(SBP ~ age, data = hers)
summary(fit_sbp_age)$coefficients["age", ]
confint(fit_sbp_age)["age", ]
boot::boot.ci(boot_result, type = c("norm", "perc", "bca"))
```

All three intervals are similar here.
When the bootstrap distribution is skewed,
the intervals differ more,
and the BCa interval, which corrects for skewness,
is the better choice [@efron1993introduction, Chapter 14].
:::
