The HERS data are distributed with @vittinghoff2e
on the book's companion website,
as a Stata file that R can read directly:

```{r}
#| label: load-hers-url
#| eval: false
#| code-fold: false
# one unbroken string, so that link checkers test the whole URL:
url <- "https://regression.ucsf.edu/sites/g/files/tkssra16191/files/wysiwyg/home/data/hersdata.dta" # nolint: line_length_linter.
hers <- haven::read_dta(url)
```

The `rmb` R package includes the same file,
which these notes use so that rendering does not depend on the website.
`haven::as_factor()` converts the Stata value labels to factors:

```{r}
#| label: load-hers-local
#| code-fold: false
hers <- rmb::hers |> haven::as_factor()
hers |> head()
```

The examples on this page use these variables:

| Variable | Meaning |
|----------|---------|
| `HT` | Randomized treatment: placebo or hormone therapy |
| `age` | Age at baseline (years) |
| `raceth` | Race/ethnicity: White, African American, or Other |
| `exercise` | Exercises at least three times per week (no/yes) |
| `BMI` | Body mass index at baseline (kg/m^2^) |
| `SBP` | Systolic blood pressure at baseline (mmHg) |
| `glucose` | Fasting glucose at baseline (mg/dL) |
| `glucose1` | Fasting glucose at the year-1 visit (mg/dL) |

```{r}
#| label: glimpse-hers
#| code-fold: false
hers |>
  dplyr::select(HT, age, raceth, exercise, BMI, SBP, glucose, glucose1) |>
  dplyr::glimpse()
```
