It is numerically better to compute the log-likelihood first
and exponentiate it to get the likelihood,
because a product of 100 densities can underflow to zero:

```{r}
#| code-fold: false
loglik <- function(mu, sigma, x) {
  n <- length(x)
  normalizing_constant <- -n / 2 * log(2 * pi * sigma^2)
  # written with sum(x), sum(x^2) so that `mu` can be a vector of values:
  kernel <- -1 / (2 * sigma^2) * (sum(x^2) - 2 * sum(x) * mu + n * mu^2)
  normalizing_constant + kernel
}

lik <- function(...) exp(loglik(...))

# log-likelihood at the MLEs:
loglik(mu = mu_hat, sigma = sigma_hat, x = glucose_data)
```

{{< slidebreak >}}

@fig-hers-lik-mu graphs the likelihood and log-likelihood
as functions of $\mu$,
with $\sigma$ fixed at $\hat\sigma_{\text{ML}}$.

::: {#fig-hers-lik-mu layout-ncol="2"}

::: {#fig-likelihood-mu}

```{r}
ggplot2::ggplot() +
  ggplot2::geom_function(
    fun = lik,
    args = list(sigma = sigma_hat, x = glucose_data)
  ) +
  ggplot2::xlim(mu_hat + c(-1, 1) * sigma_hat) +
  ggplot2::xlab("mu") +
  ggplot2::ylab("likelihood") +
  ggplot2::geom_vline(xintercept = mu_hat, col = "red")
```

Likelihood

:::

::: {#fig-loglik-mu}

```{r}
ggplot2::ggplot() +
  ggplot2::geom_function(
    fun = loglik,
    args = list(sigma = sigma_hat, x = glucose_data)
  ) +
  ggplot2::xlim(mu_hat + c(-1, 1) * sigma_hat) +
  ggplot2::xlab("mu") +
  ggplot2::ylab("log-likelihood") +
  ggplot2::geom_vline(xintercept = mu_hat, col = "red")
```

Log-likelihood

:::

Likelihood and log-likelihood of the HERS glucose data as functions of $\mu$,
with $\sigma = \hat\sigma_{\text{ML}}$; the red line marks $\hat\mu_{\text{ML}}$

:::

{{< slidebreak >}}

@fig-hers-lik-sigma graphs them as functions of $\sigma$,
with $\mu$ fixed at $\hat\mu_{\text{ML}}$.

::: {#fig-hers-lik-sigma layout-ncol="2"}

::: {#fig-hers-lik-sigma-lik}

```{r}
ggplot2::ggplot() +
  ggplot2::geom_function(
    fun = lik,
    args = list(mu = mu_hat, x = glucose_data)
  ) +
  ggplot2::xlim(sigma_hat * c(0.9, 1.1)) +
  ggplot2::geom_vline(xintercept = sigma_hat, col = "red") +
  ggplot2::xlab("sigma") +
  ggplot2::ylab("likelihood")
```

Likelihood

:::

::: {#fig-hers-lik-sigma-llik}

```{r}
ggplot2::ggplot() +
  ggplot2::geom_function(
    fun = loglik,
    args = list(mu = mu_hat, x = glucose_data)
  ) +
  ggplot2::xlim(sigma_hat * c(0.9, 1.1)) +
  ggplot2::geom_vline(xintercept = sigma_hat, col = "red") +
  ggplot2::xlab("sigma") +
  ggplot2::ylab("log-likelihood")
```

Log-likelihood

:::

Likelihood and log-likelihood of the HERS glucose data as functions of $\sigma$,
with $\mu = \hat\mu_{\text{ML}}$; the red line marks $\hat\sigma_{\text{ML}}$

:::
