{{< include sds/_subfiles/shared/_sec_hers_intro.qmd >}}

We model the distribution of fasting glucose
among HERS participants who do not have diabetes and do not exercise.

The HERS data are distributed with @vittinghoff2e
on the book's companion website:

```{r}
#| eval: false
# one unbroken string, so that link checkers test the whole URL:
url <- "https://regression.ucsf.edu/sites/g/files/tkssra6706/f/wysiwyg/home/data/hersdata.dta" # nolint: line_length_linter.
hers <- haven::read_dta(url)
```

The `rmb` R package includes the same file,
which these notes use so that rendering does not depend on the website
(@tbl-HERS):

```{r}
hers <- rmb::hers |> haven::zap_labels()
```

::: {#tbl-HERS}

```{r}
hers |> head()
```

The first rows of the HERS dataset

:::

{{< slidebreak >}}

To keep the likelihood graphs readable,
we use only the first 100 eligible participants
(@fig-hers-glucose-hist);
with the whole subset, the likelihood would be too concentrated
to graph clearly.

```{r}
n_obs <- 100

data1 <-
  hers |>
  dplyr::filter(
    diabetes == 0,
    exercise == 0
  ) |>
  head(n_obs)

glucose_data <- data1$glucose
```

::: {#fig-hers-glucose-hist}

```{r}
plot1 <-
  data1 |>
  ggplot2::ggplot() +
  ggplot2::aes(x = glucose) +
  ggplot2::geom_histogram(
    ggplot2::aes(y = ggplot2::after_stat(density)),
    bins = 20
  ) +
  ggplot2::xlab("Fasting glucose (mg/dL)")

print(plot1)
```

Fasting glucose among 100 HERS participants without diabetes who do not exercise

:::

The histogram is irregular, as histograms of 100 observations often are,
with a somewhat longer right tail than left tail.
A Gaussian model is a rough but usable starting point.
