Suppose we test the null hypothesis $H_0: \mu = \mu_0$,
with $\mu_0 = 95$ mg/dL,
at significance level $\alpha = 0.05$,
and suppose for simplicity that $\sigma$ is known,
equal to $\hat\sigma_{\text{ML}}$.
Then under $H_0$, $\bar X \sim \ndist{\mu_0, \sigma^2/n}$,
and the test rejects $H_0$ when $\bar x$ falls outside the
non-rejection interval

$$\mu_0 \pm z_{1 - \alpha/2} \frac{\sigma}{\sqrt{n}}$$

```{r}
mu0 <- 95
se <- se_mu_hat(n = n_obs, sigma = sigma_hat)
margin <- qnorm(0.975) * se
bounds <- mu0 + c(lower = -1, upper = 1) * margin
bounds
```

{{< slidebreak >}}

{{< include sds/_subfiles/HERS-example/_def-power.qmd >}}

{{< slidebreak >}}

For this test, under $\mu = \mu_1$, $\bar X \sim \ndist{\mu_1, \sigma^2/n}$, so:

$$
\text{power}(\mu_1) =
\Phi\paren{\frac{\mu_0 - z_{1-\alpha/2}\,\sigma/\sqrt{n} - \mu_1}{\sigma/\sqrt{n}}}
+ 1 - \Phi\paren{\frac{\mu_0 + z_{1-\alpha/2}\,\sigma/\sqrt{n} - \mu_1}{\sigma/\sqrt{n}}}
$$

where $\Phi$ is the standard Gaussian CDF.
For example, the power against $\mu_1 = 100$ mg/dL is:

```{r}
power <- function(n, null, alt, sigma) {
  se <- sigma / sqrt(n)
  lower <- null - qnorm(0.975) * se
  upper <- null + qnorm(0.975) * se
  pnorm((lower - alt) / se) + pnorm((upper - alt) / se, lower.tail = FALSE)
}
mu1 <- 100
power(n = n_obs, null = mu0, alt = mu1, sigma = sigma_hat)
```

{{< slidebreak >}}

@fig-hers-power graphs the power as a function of the sample size.

::: {#fig-hers-power}

```{r}
ggplot2::ggplot() +
  ggplot2::geom_function(
    fun = power,
    args = list(null = mu0, alt = mu1, sigma = sigma_hat),
    n = 199
  ) +
  ggplot2::xlim(c(2, 200)) +
  ggplot2::ylim(0, 1) +
  ggplot2::ylab("Power") +
  ggplot2::xlab("n")
```

Power of the test of $H_0: \mu = 95$ against $\mu_1 = 100$ mg/dL,
by sample size

:::

::: notes
The alternative $\mu_1$ should be chosen before seeing the data,
as a difference worth detecting.
Power computed at $\mu_1 = \hat\mu$ ("observed power")
is a function of the p-value,
so it adds no information about the data already analyzed
[@hoenig2001abuse].
:::
