To check how maximum likelihood estimation behaves for this model,
we simulate many datasets from a Gaussian distribution
whose parameters equal the HERS estimates,
analyze each one,
and summarize the results.

`do_one_sim()` simulates and analyzes one dataset:
it computes $\hat\mu$, its estimated standard error,
a 95% $t$-based confidence interval for $\mu$,
and the $t$-test of $H_0: \mu = \mu_0$.

```{r}
do_one_sim <- function(
  n = 100,
  mu = mu_hat,
  mu_0 = mu_hat * 0.9,
  sigma2 = sigma_sq_hat,
  return_data = FALSE
) {
  # generate data
  x <- rnorm(n = n, mean = mu, sd = sqrt(sigma2))

  # analyze data
  est <- mean(x)
  se_est <- sd(x) / sqrt(n)
  confint <- est + c(-1, 1) * se_est * qt(0.975, df = n - 1)
  tstat <- abs(est - mu_0) / se_est
  pval <- 2 * pt(q = tstat, df = n - 1, lower.tail = FALSE)

  results <- tibble::tibble(
    mu_hat = est,
    sigma_hat = sd(x),
    se_hat = se_est,
    confint_left = confint[1],
    confint_right = confint[2],
    tstat = tstat,
    pval = pval,
    confint_covers = dplyr::between(mu, confint[1], confint[2]),
    test_rejects = pval < 0.05
  )

  if (return_data) {
    list(data = x, results = results)
  } else {
    results
  }
}
```

{{< slidebreak >}}

To check `do_one_sim()`,
we compare its output with `stats::t.test()` on the same simulated data:

```{r}
set.seed(1)
sim_output <- do_one_sim(mu_0 = 80, return_data = TRUE)
t_test <- t.test(sim_output$data, mu = 80)

dplyr::bind_rows(
  `do_one_sim()` = sim_output$results |>
    dplyr::select(mu_hat, se_hat, confint_left, confint_right, pval),
  `t.test()` = tibble::tibble(
    mu_hat = unname(t_test$estimate),
    se_hat = t_test$stderr,
    confint_left = t_test$conf.int[1],
    confint_right = t_test$conf.int[2],
    pval = t_test$p.value
  ),
  .id = "source"
)
```

The two rows agree.

{{< slidebreak >}}

`do_n_sims()` repeats the simulation `n_sims` times,
with a different random seed for each dataset:

```{r}
do_n_sims <- function(n_sims = 1000, ...) {
  lapply(seq_len(n_sims), function(i) {
    set.seed(i)
    do_one_sim(...) |>
      dplyr::mutate(sim_number = i, .before = dplyr::everything())
  }) |>
    dplyr::bind_rows()
}

sim_results <- do_n_sims(n_sims = 1000, n = 100)
sim_results
```

{{< slidebreak >}}

`summarize_sim()` compares the simulation results
with the true data-generating parameters:

```{r}
summarize_sim <- function(sim_results, mu = mu_hat, sigma2 = sigma_sq_hat, n = 100) {
  true_se <- sqrt(sigma2 / n)
  sim_results |>
    dplyr::summarize(
      bias_mu_hat = mean(mu_hat) - mu,
      sd_mu_hat = sd(mu_hat),
      true_se = true_se,
      bias_se_hat = mean(se_hat) - true_se,
      coverage = mean(confint_covers),
      power = mean(test_rejects)
    )
}

sim_summary <- summarize_sim(sim_results)
sim_summary
```

Across `r nrow(sim_results)` simulated datasets:

- the average error of $\hat\mu$ is
  `r signif(sim_summary$bias_mu_hat, 2)` mg/dL,
  small relative to its standard error of
  `r signif(sim_summary$true_se, 3)` mg/dL,
  consistent with $\hat\mu$ being unbiased;
- the standard deviation of the $\hat\mu$ values,
  `r signif(sim_summary$sd_mu_hat, 3)`,
  is close to the true standard error;
- the 95% confidence intervals covered the true $\mu$ in
  `r round(100 * sim_summary$coverage, 1)`% of datasets,
  close to their nominal 95%;
- the test of $H_0: \mu = 0.9\,\hat\mu$ rejected in
  `r round(100 * sim_summary$power, 1)`% of datasets:
  with $n = 100$, a 10% difference in the mean is easy to detect.

Changing the sample size, the true $\mu$, or $\sigma^2$
in `do_n_sims()` shows how these properties depend on them.
