:::{#exm-dobson-cyclones-Newton}
#### Finding the MLE using the Newton-Raphson algorithm

We found the MLE $\hat{\lambda} = \bar{x}$
by solving the score equation $\score(\lambda) = 0$ algebraically
(@exr-solve-score-equation).
If we could not have solved it,
we could instead start from an initial guess such as $\esttmpl = 3$
and apply the [Newton-Raphson algorithm](#sec-newton-raphson).

```{r}
#| label: initial-guess
#| code-fold: false
cur_lambda_est <- 3
cur_lambda_est
```

:::

{{< slidebreak >}}

From @exr-cyclone-score-fn and @exr-hessian,
the score function and Hessian are:

$$
\begin{aligned}
\score(\lambda; \vx) &= \frac{`r sumx`}{\lambda} - `r n`\\
\hessian(\lambda; \vx) &= -\frac{`r sumx`}{\lambda^2}
\end{aligned}
$$

So the first-order Taylor approximation of the score function
around $\esttmpl$ is:

$$
\begin{aligned}
\score(\lambda)
&\approx \score^*(\lambda)\\
&\eqdef \score(\esttmpl) + \hessian(\esttmpl)(\lambda - \esttmpl)\\
&= \paren{\frac{`r sumx`}{\esttmpl} - `r n`}
 + \paren{-\frac{`r sumx`}{\sqf{\esttmpl}}} (\lambda - \esttmpl)
\end{aligned}
$$

{{< slidebreak >}}

@fig-cyclone-newton-step1 compares the score function
and the approximate score function at $\esttmpl = `r cur_lambda_est`$.

:::{#fig-cyclone-newton-step1}

```{r}
# score(), hessian() and loglik() are defined earlier on the page
# nolint start: object_usage_linter.
approx_score <- function(lambda, lhat, ...) {
  score(lambda = lhat, ...) +
    hessian(lambda = lhat, ...) * (lambda - lhat)
}
# nolint end

point_size <- 5

plot1 <- ggplot2::ggplot() +
  ggplot2::geom_function(
    fun = score,
    ggplot2::aes(col = "score function"),
    n = 1001
  ) +
  ggplot2::geom_function(
    fun = approx_score,
    ggplot2::aes(col = "approximate score function"),
    n = 1001,
    args = list(lhat = cur_lambda_est)
  ) +
  ggplot2::geom_point(
    size = point_size,
    ggplot2::aes(
      x = cur_lambda_est, y = score(lambda = cur_lambda_est),
      col = "current estimate"
    )
  ) +
  ggplot2::geom_point(
    size = point_size,
    ggplot2::aes(x = xbar, y = 0, col = "MLE")
  ) +
  ggplot2::xlim(min(cyclones$number), max(cyclones$number)) +
  ggplot2::ylab("l'(lambda)") +
  ggplot2::xlab("lambda") +
  ggplot2::geom_hline(yintercept = 0)

print(plot1)
```

Score function of the cyclone data and its first-order approximation
at the initial guess

:::

{{< slidebreak >}}

Approximating the score function by a linear function
is equivalent to approximating the log-likelihood
by a second-order Taylor polynomial
(@fig-cyclone-newton-step1-loglik):

$$
\llik^*(\lambda) \eqdef
\llik(\esttmpl) +
(\lambda - \esttmpl) \score(\esttmpl) +
\frac{1}{2}\hessian(\esttmpl)(\lambda - \esttmpl)^2
$$

:::{#fig-cyclone-newton-step1-loglik}

```{r}
# nolint start: object_usage_linter.
approx_loglik <- function(lambda, lhat, ...) {
  loglik(lambda = lhat, ...) +
    score(lambda = lhat, ...) * (lambda - lhat) +
    1 / 2 * hessian(lambda = lhat, ...) * (lambda - lhat)^2
}
# nolint end

plot_loglik <- ggplot2::ggplot() +
  ggplot2::geom_function(
    fun = loglik,
    ggplot2::aes(col = "log-likelihood"),
    n = 1001
  ) +
  ggplot2::geom_function(
    fun = approx_loglik,
    ggplot2::aes(col = "approximate log-likelihood"),
    n = 1001,
    args = list(lhat = cur_lambda_est)
  ) +
  ggplot2::geom_point(
    size = point_size,
    ggplot2::aes(
      x = cur_lambda_est, y = loglik(lambda = cur_lambda_est),
      col = "current estimate"
    )
  ) +
  ggplot2::geom_point(
    size = point_size,
    ggplot2::aes(x = xbar, y = loglik(xbar), col = "MLE")
  ) +
  ggplot2::xlim(min(cyclones$number) - 1, max(cyclones$number)) +
  ggplot2::ylab("log-likelihood") +
  ggplot2::xlab("lambda")

print(plot_loglik)
```

Log-likelihood of the cyclone data and its second-order approximation
at the initial guess

:::

{{< slidebreak >}}

Solving the approximate score equation $\score^*(\lambda) = 0$
gives the next estimate:

$$
\begin{aligned}
\lambda
&= \esttmpl - \score(\esttmpl) \cd \inv{\hessian(\esttmpl)}\\
&= `r round(cur_lambda_est - score(cur_lambda_est) / hessian(cur_lambda_est), 4)`
\end{aligned}
$$

```{r}
#| label: NR-one-step
#| code-fold: false
new_lambda_est <-
  cur_lambda_est - score(cur_lambda_est) / hessian(cur_lambda_est)
new_lambda_est
```

:::{#fig-cyclone-newton-step1-with-new-est}

```{r}
plot2 <- plot1 +
  ggplot2::geom_point(
    size = point_size,
    ggplot2::aes(x = new_lambda_est, y = 0, col = "new estimate")
  ) +
  ggplot2::geom_segment(
    arrow = grid::arrow(),
    linewidth = 2,
    alpha = .7,
    ggplot2::aes(
      x = cur_lambda_est,
      y = approx_score(lhat = cur_lambda_est, lambda = cur_lambda_est),
      xend = new_lambda_est,
      yend = 0,
      col = "update"
    )
  )
print(plot2)
```

The first Newton-Raphson update:
the new estimate is where the approximate score function crosses zero

:::

{{< slidebreak >}}

We update $\esttmpl \leftarrow `r round(new_lambda_est, 4)`$
and repeat the process
(@fig-cyclone-newton-step2).

:::{#fig-cyclone-newton-step2}

```{r}
plot2 +
  ggplot2::geom_function(
    fun = approx_score,
    ggplot2::aes(col = "new approximate score function"),
    n = 1001,
    args = list(lhat = new_lambda_est)
  ) +
  ggplot2::geom_point(
    size = point_size,
    ggplot2::aes(
      x = new_lambda_est, y = score(lambda = new_lambda_est),
      col = "new estimate"
    )
  )
```

The approximate score function at the updated estimate

:::

{{< slidebreak >}}

We repeat this process until the log-likelihood stops changing
(@tbl-mle-converge).

:::{#tbl-mle-converge}

```{r}
#| code-fold: false
cur_lambda_est <- 3 # restart from the initial guess
tolerance <- 10^-4
max_iter <- 100
nr_info <- tibble::tibble(
  iteration = 0,
  lambda = cur_lambda_est,
  `log(likelihood)` = loglik(cur_lambda_est),
  score = score(cur_lambda_est),
  hessian = hessian(cur_lambda_est)
)

for (cur_iter in 1:max_iter) {
  new_lambda_est <-
    cur_lambda_est - score(cur_lambda_est) / hessian(cur_lambda_est)

  diff_loglik <- loglik(new_lambda_est) - loglik(cur_lambda_est)

  nr_info <- nr_info |>
    dplyr::bind_rows(
      tibble::tibble(
        iteration = cur_iter,
        lambda = new_lambda_est,
        `log(likelihood)` = loglik(new_lambda_est),
        score = score(new_lambda_est),
        hessian = hessian(new_lambda_est),
        `diff(loglik)` = diff_loglik
      )
    )

  cur_lambda_est <- new_lambda_est

  if (abs(diff_loglik) < tolerance) {
    break
  }
}

nr_info |> knitr::kable(digits = 5)
```

Convergence of the Newton-Raphson algorithm to the MLE for the cyclone data

:::

The final estimate matches the closed-form MLE,
$\bar x = `r round(xbar, 5)`$ (@exr-find-mle).

{{< slidebreak >}}

:::{#fig-NR-converge}

```{r}
ll_plot +
  ggplot2::geom_segment(
    data = nr_info,
    arrow = grid::arrow(),
    alpha = .7,
    ggplot2::aes(
      x = lambda,
      xend = dplyr::lead(lambda),
      y = `log(likelihood)`,
      yend = dplyr::lead(`log(likelihood)`),
      col = factor(iteration)
    )
  ) +
  ggplot2::labs(col = "iteration")
```

Newton-Raphson steps toward the MLE of the Poisson model (@eq-iid-model)
for the cyclone data

:::
