Suppose we want to learn how many cyclones to expect per season.

::: {#fig-dobson-cyclone-time-series}

```{r}
cyclones |>
  dplyr::mutate(years = factor(years, levels = years)) |>
  ggplot2::ggplot() +
  ggplot2::aes(x = years, y = number, group = 1) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::xlab("Season") +
  ggplot2::ylab("Number of cyclones") +
  ggplot2::expand_limits(y = 0) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(vjust = .5, angle = 45))
```

Number of tropical cyclones per season in northeastern Australia,
1956/57 to 1968/69

:::

@fig-dobson-cyclone-time-series shows no obvious trend,
and no obvious correlation between adjacent seasons,
so we assume that the seasons' counts are mutually independent.
We also assume that they are identically distributed,
with a common distribution $\Pr(X = x)$;
the expression has no index $i$,
because the distribution is the same for every season.

{{< slidebreak >}}

@fig-cyclones-bar-plot shows the empirical distribution of the counts.

::: {#fig-cyclones-bar-plot}

```{r}
cyclones |>
  ggplot2::ggplot() +
  ggplot2::aes(x = number) +
  ggplot2::geom_bar() +
  ggplot2::expand_limits(x = 0) +
  ggplot2::xlab("Number of cyclones") +
  ggplot2::ylab("Count (number of seasons)")
```

Bar plot of cyclones per season

:::

{{< slidebreak >}}

@tbl-dobson-cyclones-sumstat provides summary statistics.

::: {#tbl-dobson-cyclones-sumstat}

```{r}
n <- nrow(cyclones)
sumx <- sum(cyclones$number)
xbar <- mean(cyclones$number)
tibble::tibble(
  seasons = n,
  total = sumx,
  mean = xbar,
  variance = var(cyclones$number)
) |>
  knitr::kable(digits = 2)
```

Summary statistics for the `cyclones` data

:::
