:::{#exm-kfold-simulated}
#### Cross-validation tracks the generalization error

This example applies [$k$-fold cross-validation](#def-kfold)
with $k = 5$
to the simulated training set of @exm-train-test-error-simulated,
using only those 30 observations.
The function `cv_mse()` computes @eq-cv-estimate:
it assigns each observation to a fold,
fits the model once per fold without that fold,
and averages the squared prediction errors.

```{r}
#| label: cv-mse-function
#| code-fold: false
assign_folds <- function(n, k) {
  sample(rep_len(seq_len(k), n))
}

cv_mse <- function(formula, data, folds) {
  y <- stats::model.response(stats::model.frame(formula, data))
  sq_errors <- numeric(nrow(data))
  for (j in unique(folds)) {
    in_fold <- folds == j
    fit <- lm(formula, data = data[!in_fold, ])
    pred <- predict(fit, newdata = data[in_fold, ])
    sq_errors[in_fold] <- (pred - y[in_fold])^2
  }
  mean(sq_errors)
}

set.seed(5)
sim_folds <- assign_folds(nrow(sim_train), k = 5)
table(fold = sim_folds)
```

The same fold assignment is used for every degree,
so the degrees are compared on the same splits.

```{r}
#| label: cv-simulated-table
sim_errors <- sim_errors |>
  dplyr::mutate(
    cv_5 = purrr::map_dbl(
      degree,
      \(d) cv_mse(y ~ poly(x, d), sim_train, sim_folds)
    )
  )

sim_errors
```

:::{#fig-cv-simulated}

```{r}
#| label: cv-simulated-plot
#| fig-alt: Mean squared error against polynomial degree from 1 to 10,
#|   with three curves for the training error,
#|   the estimated generalization error,
#|   and the 5-fold cross-validation estimate.
#|   The cross-validation curve drops at degree 3,
#|   as the generalization error curve does,
#|   but it is noisier and rises steeply at the highest degrees,
#|   while the training error keeps falling.
sim_errors |>
  tidyr::pivot_longer(
    c(training, generalization, cv_5),
    names_to = "error",
    values_to = "mse"
  ) |>
  dplyr::mutate(
    error = factor(
      error,
      levels = c("training", "cv_5", "generalization"),
      labels = c("training", "5-fold CV", "generalization")
    )
  ) |>
  ggplot2::ggplot(ggplot2::aes(degree, mse, colour = error, shape = error)) +
  ggplot2::geom_hline(yintercept = sim_sigma^2, linetype = "dashed") +
  ggplot2::geom_line() +
  ggplot2::geom_point(size = 2) +
  ggplot2::scale_x_continuous(breaks = sim_degrees) +
  ggplot2::scale_y_log10() +
  ggplot2::labs(
    x = "polynomial degree",
    y = "mean squared error (log scale)",
    colour = NULL,
    shape = NULL
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "top")
```

@fig-train-test-error with the 5-fold cross-validation estimate added.
Only the training and cross-validation curves
could be computed from real data.
:::

```{r}
#| label: cv-simulated-summary
sim_best_cv <- sim_errors$degree[which.min(sim_errors$cv_5)]
c(
  best_by_cv = sim_best_cv,
  best_by_generalization = sim_best_generalization
)
```

The cross-validation estimate is smallest at degree `r sim_best_cv`,
and the generalization error at degree `r sim_best_generalization`.
Unlike the training error,
the cross-validation estimate rises again for the most flexible fits,
as the generalization error does.
It also jumps around more from one degree to the next,
because it rests on only 30 observations
rather than on $10{,}000$ new ones.

:::
