:::{#exm-one-se-rule}
#### The one-standard-error rule for the simulated polynomials

This example applies 10-fold cross-validation
to the simulated training set of @exm-train-test-error-simulated,
records the mean squared error in each fold,
and applies @def-one-se-rule to the ten degrees.

```{r}
#| label: one-se-simulated
cv_fold_mse <- function(formula, data, folds) {
  y <- stats::model.response(stats::model.frame(formula, data))
  purrr::map_dbl(sort(unique(folds)), \(j) {
    in_fold <- folds == j
    fit <- lm(formula, data = data[!in_fold, ])
    mean((predict(fit, newdata = data[in_fold, ]) - y[in_fold])^2)
  })
}

set.seed(10)
sim_folds_ten <- assign_folds(nrow(sim_train), k = 10)

sim_cv_ten <- tibble::tibble(degree = sim_degrees) |>
  dplyr::mutate(
    fold_mse = purrr::map(
      degree,
      \(d) cv_fold_mse(y ~ poly(x, d), sim_train, sim_folds_ten)
    ),
    cv = purrr::map_dbl(fold_mse, mean),
    se = purrr::map_dbl(fold_mse, \(v) sd(v) / sqrt(length(v)))
  ) |>
  dplyr::select(-fold_mse)

sim_min <- dplyr::slice_min(sim_cv_ten, cv, n = 1)
sim_one_se <-
  sim_cv_ten |>
  dplyr::filter(cv <= sim_min$cv + sim_min$se) |>
  dplyr::slice_min(degree, n = 1)

sim_cv_ten
```

With 30 observations and 10 folds,
each fold holds 3 observations,
so all ten folds have equal size
and `cv` is $\cvest{10}$ (@rem-cv-estimate).

:::{#fig-one-se-rule}

```{r}
#| label: one-se-rule-plot
#| fig-alt: Ten-fold cross-validation estimate against polynomial degree
#|   from 1 to 10, with error bars of one standard error.
#|   A dashed horizontal line sits one standard error above the
#|   smallest estimate.
sim_cv_ten |>
  ggplot2::ggplot(ggplot2::aes(degree, cv)) +
  ggplot2::geom_hline(
    yintercept = sim_min$cv + sim_min$se,
    linetype = "dashed"
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = cv - se, ymax = cv + se),
    width = 0.2
  ) +
  ggplot2::geom_point(size = 2) +
  ggplot2::scale_x_continuous(breaks = sim_degrees) +
  ggplot2::scale_y_log10() +
  ggplot2::labs(
    x = "polynomial degree",
    y = "10-fold CV estimate (log scale)"
  ) +
  ggplot2::theme_bw()
```

10-fold cross-validation estimates,
with bars of one standard error (@eq-cv-se).
The dashed line is one standard error above the smallest estimate.
:::

The smallest estimate is at degree `r sim_min$degree`.
The least flexible degree within one standard error of it
is degree `r sim_one_se$degree`,
which is the one-standard-error rule's choice.
The rule prefers the simpler model
when the data cannot clearly tell the two apart.
Because the data are simulated,
we can check both choices against
the estimated generalization errors of @exm-train-test-error-simulated:
`r round(sim_errors$generalization[sim_min$degree], 3)` for degree `r sim_min$degree`
and `r round(sim_errors$generalization[sim_one_se$degree], 3)` for degree `r sim_one_se$degree`.

:::
