<!--
This fragment's R chunks use objects made by earlier fragments
of model-validation.qmd, so it renders only after them:
assign_folds() and cv_squared_errors(), from _exm-kfold-simulated.qmd,
and sim_train, sim_degrees and sim_errors,
from _exm-train-test-error-simulated.qmd.
-->
:::{#exm-one-se-rule}
#### The one-standard-error rule for the simulated polynomials

This example applies 10-fold cross-validation
to the simulated training set of @exm-train-test-error-simulated,
records the mean squared error in each fold,
and applies @def-one-se-rule to the ten degrees.

```{r}
#| label: one-se-simulated
set.seed(10)
sim_folds_ten <- assign_folds(nrow(sim_train), k = 10)

sim_cv_ten <- tibble::tibble(degree = sim_degrees) |>
  dplyr::mutate(
    fold_mse = purrr::map(
      degree,
      \(d) {
        sq_errors <- cv_squared_errors(y ~ poly(x, d), sim_train, sim_folds_ten)
        as.numeric(tapply(sq_errors, sim_folds_ten, mean))
      }
    ),
    cv = purrr::map_dbl(fold_mse, mean),
    se = purrr::map_dbl(fold_mse, \(v) sd(v) / sqrt(length(v)))
  ) |>
  dplyr::select(-fold_mse)

sim_min <- dplyr::slice_min(sim_cv_ten, cv, n = 1)
sim_one_se <-
  sim_cv_ten |>
  dplyr::filter(cv <= sim_min$cv + sim_min$se) |>
  dplyr::slice_min(degree, n = 1)

sim_cv_ten
```

With 30 observations and 10 folds,
each fold holds 3 observations,
so all ten folds have equal size
and `cv` is $\cvest{10}$ (@rem-cv-estimate).

:::{#fig-one-se-rule}

```{r}
#| label: one-se-rule-plot
#| fig-alt: Ten-fold cross-validation estimate against polynomial degree
#|   from 1 to 10, with error bars of one fold-based standard error.
#|   A dashed horizontal line sits one fold-based standard error
#|   above the smallest estimate.
sim_cv_ten |>
  ggplot2::ggplot(ggplot2::aes(degree, cv)) +
  ggplot2::geom_hline(
    yintercept = sim_min$cv + sim_min$se,
    linetype = "dashed"
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = cv - se, ymax = cv + se),
    width = 0.2
  ) +
  ggplot2::geom_point(size = 2) +
  ggplot2::scale_x_continuous(breaks = sim_degrees) +
  ggplot2::scale_y_log10() +
  ggplot2::labs(
    x = "polynomial degree",
    y = "10-fold CV estimate (log scale)"
  ) +
  ggplot2::theme_bw()
```

10-fold cross-validation estimates,
with bars of one [fold-based standard error](#def-cv-se) (@eq-cv-se).
The dashed line is one fold-based standard error above the smallest estimate.
:::

The smallest estimate is at degree `r sim_min$degree`.
The lowest degree within one fold-based standard error of it
is degree `r sim_one_se$degree`,
which is the one-standard-error rule's choice.
The rule prefers the simpler model
when the data cannot clearly tell it apart from the degree with the smallest estimate.
Because the data are simulated,
we can check both choices against
the estimated generalization errors of @exm-train-test-error-simulated:
`r round(sim_errors$generalization[sim_min$degree], 3)` for degree `r sim_min$degree`
and `r round(sim_errors$generalization[sim_one_se$degree], 3)` for degree `r sim_one_se$degree`.

:::
