<!--
This fragment's R chunks use objects made by earlier fragments
of model-validation.qmd, so it renders only after them:
mtcars_degrees, from _exm-kfold-mtcars.qmd.
-->
:::{#exm-loocv-shortcut}
#### Leave-one-out cross-validation of the `mtcars` polynomials

This example computes $\cvest{32}$ for the four `mtcars` polynomials
of @exr-overfitting in two ways:
by fitting each model 32 times, as in @eq-loocv,
and by the single-fit formula of @thm-loocv-shortcut,
with the leverages from `hatvalues()`.

```{r}
#| label: loocv-mtcars
#| code-fold: false
loocv_by_refitting <- function(formula, data) {
  y <- stats::model.response(stats::model.frame(formula, data))
  sq_errors <- purrr::map_dbl(seq_len(nrow(data)), \(i) {
    fit <- lm(formula, data = data[-i, ])
    (predict(fit, newdata = data[i, ]) - y[i])^2
  })
  mean(sq_errors)
}

loocv_by_formula <- function(formula, data) {
  fit <- lm(formula, data = data)
  pred_error <- fitted(fit) - stats::model.response(model.frame(fit))
  mean((pred_error / (1 - hatvalues(fit)))^2)
}

mtcars_loocv <- tibble::tibble(degree = mtcars_degrees) |>
  dplyr::mutate(
    refitting = purrr::map_dbl(
      degree,
      \(d) loocv_by_refitting(mpg ~ poly(wt, d), mtcars)
    ),
    formula = purrr::map_dbl(
      degree,
      \(d) loocv_by_formula(mpg ~ poly(wt, d), mtcars)
    ),
    loocv_rmse = sqrt(formula)
  )

mtcars_loocv
```

```{r}
#| label: loocv-mtcars-check
loocv_max_diff <- max(abs(mtcars_loocv$refitting - mtcars_loocv$formula))
mtcars_best_loocv <-
  mtcars_loocv$degree[which.min(mtcars_loocv$formula)]
c(max_difference = loocv_max_diff, best_degree = mtcars_best_loocv)
```

The two columns differ by at most `r format(signif(loocv_max_diff, 2), scientific = TRUE)`,
which is rounding error.
The single-fit formula needs one fit per model instead of 32.
LOOCV picks degree `r mtcars_best_loocv`,
and because LOOCV is not random (@sol-loocv),
rerunning it always gives the same choice.

:::
