<!--
This fragment's R chunks use objects made by earlier fragments
of model-validation.qmd, so it renders only after them:
sim_train, sim_new, sim_errors, sim_sigma, sim_degrees,
sim_draw(), mse() and sim_best_generalization,
from the chunks of _exm-train-test-error-simulated.qmd.
-->
:::{#exm-kfold-simulated}
#### Cross-validation versus the generalization error

This example applies [$k$-fold cross-validation](#def-kfold)
with $k = 5$
to the simulated training set of @exm-train-test-error-simulated,
using only those 30 observations.
The function `assign_folds()` assigns each observation to a fold at random,
and `cv_mse()` computes @eq-cv-estimate for a given assignment:
it fits the model once per fold without that fold
and averages the squared prediction errors.

```{r}
#| label: cv-mse-function
#| code-fold: false
assign_folds <- function(n, k) {
  sample(rep_len(seq_len(k), n))
}

cv_mse <- function(formula, data, folds) {
  y <- stats::model.response(stats::model.frame(formula, data))
  sq_errors <- numeric(nrow(data))
  for (j in unique(folds)) {
    in_fold <- folds == j
    fit <- lm(formula, data = data[!in_fold, ])
    pred <- predict(fit, newdata = data[in_fold, ])
    sq_errors[in_fold] <- (pred - y[in_fold])^2
  }
  mean(sq_errors)
}

set.seed(5)
sim_folds <- assign_folds(nrow(sim_train), k = 5)
table(fold = sim_folds)
```

The same fold assignment is used for every degree,
so the degrees are compared on the same splits.

```{r}
#| label: cv-simulated-table
sim_errors_cv <- sim_errors |>
  dplyr::mutate(
    cv_5 = purrr::map_dbl(
      degree,
      \(d) cv_mse(y ~ poly(x, d), sim_train, sim_folds)
    )
  )

sim_errors_cv
```

:::{#fig-cv-simulated}

```{r}
#| label: cv-simulated-plot
#| fig-alt: Mean squared error against polynomial degree from 1 to 10,
#|   with three curves for the training error,
#|   the estimated generalization error,
#|   and the 5-fold cross-validation estimate.
#|   The training error falls as the degree grows.
#|   The other two curves fall at the low degrees,
#|   then level off;
#|   the cross-validation curve rises sharply at the highest degrees.
sim_errors_cv |>
  tidyr::pivot_longer(
    c(training, generalization, cv_5),
    names_to = "error",
    values_to = "mse"
  ) |>
  dplyr::mutate(
    error = factor(
      error,
      levels = c("training", "cv_5", "generalization"),
      labels = c("training", "5-fold CV", "generalization")
    )
  ) |>
  ggplot2::ggplot(ggplot2::aes(degree, mse, colour = error, shape = error)) +
  ggplot2::geom_hline(yintercept = sim_sigma^2, linetype = "dashed") +
  ggplot2::geom_line() +
  ggplot2::geom_point(size = 2) +
  ggplot2::scale_x_continuous(breaks = sim_degrees) +
  ggplot2::scale_y_log10() +
  ggplot2::labs(
    x = "polynomial degree",
    y = "mean squared error (log scale)",
    colour = NULL,
    shape = NULL
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "top")
```

@fig-train-test-error with the 5-fold cross-validation estimate added.
Only the training and cross-validation curves
could be computed from real data.
:::

```{r}
#| label: cv-simulated-summary
sim_best_cv <- sim_errors_cv$degree[which.min(sim_errors_cv$cv_5)]
sim_n_fit <- nrow(sim_train) - max(table(sim_folds))
sim_top <- sim_errors_cv[sim_errors_cv$degree == max(sim_degrees), ]
sim_cv_ratio_top <- sim_top$cv_5 / sim_top$generalization
c(
  best_by_cv = sim_best_cv,
  best_by_generalization = sim_best_generalization,
  n_per_fit = sim_n_fit,
  cv_over_generalization_at_top_degree = sim_cv_ratio_top
)
```

The cross-validation estimate is smallest at degree `r sim_best_cv`,
and the estimated generalization error at degree `r sim_best_generalization`.
Unlike the training error,
the cross-validation estimate does not keep falling as the degree grows,
so it does not simply reward higher degrees.
It is not an estimate of $\generr_{\trainset}$ for the fit to all 30 observations, though
(@rem-cv-estimate).
Each of its fits uses only `r sim_n_fit` observations.
At degree `r max(sim_degrees)`,
polynomials fitted to `r sim_n_fit` points predict their held-out points
much worse than the fit to all 30 predicts new data:
the cross-validation estimate is
`r round(sim_cv_ratio_top, 1)` times the estimated generalization error.

Part of that gap is expected:
by @rem-cv-estimate,
the cross-validation estimate targets $\expgenerr{`r sim_n_fit`}$,
not the error of a fit to 30 points.
A simulation of many training sets of each size
shows how much the size alone matters at this degree.

```{r}
#| label: cv-simulated-size-effect
sim_generalization_draws <- function(
  n_train, degree, n_reps, draw, new_data, score
) {
  purrr::map_dbl(seq_len(n_reps), \(r) {
    fit <- lm(y ~ poly(x, degree), data = draw(n_train))
    score(fit, new_data)
  })
}

sim_n_reps <- 500
set.seed(24)
sim_size_effect <- tibble::tibble(
  n_train = c(sim_n_fit, nrow(sim_train))
) |>
  dplyr::mutate(
    draws = purrr::map(
      n_train,
      \(n) {
        sim_generalization_draws(
          n, max(sim_degrees), sim_n_reps,
          draw = sim_draw, new_data = sim_new, score = mse
        )
      }
    ),
    median = purrr::map_dbl(draws, stats::median),
    quantile_90 = purrr::map_dbl(draws, \(v) stats::quantile(v, 0.9)),
    mean = purrr::map_dbl(draws, mean),
    largest = purrr::map_dbl(draws, max),
    top_two_share = purrr::map_dbl(
      draws,
      \(v) sum(sort(v, decreasing = TRUE)[1:2]) / sum(v)
    )
  ) |>
  dplyr::select(-draws)

sim_size_effect
```

Over `r sim_n_reps` simulated training sets of each size,
degree-`r max(sim_degrees)` fits to `r sim_n_fit` points
have median generalization error
`r round(sim_size_effect$median[[1]], 2)`,
against `r round(sim_size_effect$median[[2]], 2)`
for fits to `r nrow(sim_train)` points,
and 90th percentiles of
`r round(sim_size_effect$quantile_90[[1]], 2)`
and `r round(sim_size_effect$quantile_90[[2]], 2)`.
The sample means are no guide to $\expgenerr{`r sim_n_fit`}$ here:
the largest single value among the fits to `r sim_n_fit` points is
`r format(signif(sim_size_effect$largest[[1]], 2), big.mark = ",", scientific = FALSE)`,
from one fit that is wildly wrong away from the data,
and the two largest values supply
`r round(100 * sim_size_effect$top_two_share[[1]])`% and
`r round(100 * sim_size_effect$top_two_share[[2]])`% of the two means.
So fits of this degree to `r sim_n_fit` points
are much worse than fits to `r nrow(sim_train)` points,
and a large cross-validation estimate here is no surprise.
How much of this particular gap comes from the training-set size,
and how much from the chance of this one data set,
a single run cannot say.
The estimate also jumps around from one degree to the next,
because it rests on only 30 observations
rather than on $10{,}000$ new ones.

:::
