<!--
This fragment's R chunks use objects made by earlier fragments
of model-validation.qmd, so it renders only after them:
prediction_mse(), from _sol-overfitting.qmd.
-->
:::{#exm-train-test-error-simulated}
#### Training error and generalization error by degree

Real data never reveal $\generr_{\trainset}$,
but a simulation can,
because it can draw as many new observations as we like.
Here the population is known:
$X \sim \operatorname{Uniform}(0, 1)$ and
$Y \mid X = x \sim \ndist{\sin(2 \pi x), \varpar}$,
with $\sdpar = 0.3$.
We draw one training set of $n = 30$ observations,
fit a polynomial in $x$ of each degree from 1 to 10 by least squares,
and score each fit on the training set
and on $10{,}000$ new observations.
By @thm-held-out-unbiased,
the second score has expectation $\generr_{\trainset}$ given the training set,
and with $10{,}000$ observations it is a precise estimate of it.

```{r}
#| label: sim-train-test-setup
#| code-fold: false
sim_sigma <- 0.3
sim_truth <- function(x) sin(2 * pi * x)
sim_draw <- function(n) {
  x <- stats::runif(n)
  tibble::tibble(x = x, y = sim_truth(x) + stats::rnorm(n, sd = sim_sigma))
}

set.seed(1)
sim_train <- sim_draw(30)
sim_new <- sim_draw(10000)
sim_degrees <- 1:10

sim_errors <- tibble::tibble(degree = sim_degrees) |>
  dplyr::mutate(
    fit = purrr::map(degree, \(d) lm(y ~ poly(x, d), data = sim_train)),
    training = purrr::map_dbl(fit, prediction_mse, data = sim_train),
    generalization = purrr::map_dbl(fit, prediction_mse, data = sim_new)
  ) |>
  dplyr::select(-fit)

sim_errors
```

:::{#fig-train-test-error}

```{r}
#| label: train-test-error-plot
#| fig-alt: Mean squared error against polynomial degree from 1 to 10.
#|   The training error falls as the degree grows.
#|   The estimated generalization error falls at the low degrees,
#|   then levels off and rises slightly.
#|   A dashed horizontal line marks the noise variance, 0.09.
sim_errors |>
  tidyr::pivot_longer(
    c(training, generalization),
    names_to = "error",
    values_to = "mse"
  ) |>
  ggplot2::ggplot(ggplot2::aes(degree, mse, colour = error, shape = error)) +
  ggplot2::geom_hline(yintercept = sim_sigma^2, linetype = "dashed") +
  ggplot2::geom_line() +
  ggplot2::geom_point(size = 2) +
  ggplot2::scale_x_continuous(breaks = sim_degrees) +
  ggplot2::scale_y_log10() +
  ggplot2::labs(
    x = "polynomial degree",
    y = "mean squared error (log scale)",
    colour = NULL,
    shape = NULL
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "top")
```

Training mean squared error and estimated generalization error
(mean squared error on $10{,}000$ new observations)
of polynomial fits to one simulated training set of $n = 30$.
The dashed line is the noise variance $\varpar$,
the [irreducible error](https://morrison-lab.github.io/pds/variance-covariance.html#def-irreducible-error).
:::

```{r}
#| label: sim-train-test-summary
sim_best_generalization <-
  sim_errors$degree[which.min(sim_errors$generalization)]
sim_best_training <- sim_errors$degree[which.min(sim_errors$training)]
sim_n_below <- c(
  training = sum(sim_errors$training < sim_sigma^2),
  generalization = sum(sim_errors$generalization < sim_sigma^2)
)
c(
  best_by_training = sim_best_training,
  best_by_generalization = sim_best_generalization,
  n_below_sigma2 = sim_n_below
)
```

The training error is smallest at degree `r sim_best_training`
and never rises as the degree grows,
because each polynomial family contains the one before it.
The estimated generalization error is smallest at degree `r sim_best_generalization`.
No prediction rule can have generalization error below $\varpar = `r sim_sigma^2`$.
Yet the training error falls below $\varpar$
at `r sim_n_below[["training"]]` of the ten degrees,
while the estimated generalization error falls below it
at `r sim_n_below[["generalization"]]` of them:
the training error is [optimistic](#rem-training-mse-optimistic).

:::
