:::{#exm-kfold-mtcars}
#### Four-fold cross-validation of the `mtcars` polynomials

This example applies the $k = 4$ folds of @exr-kfold
to the four polynomial models of @exr-overfitting,
now using all $n = 32$ cars
and the `cv_mse()` function of @exm-kfold-simulated.
The fold assignment is random,
so the estimate is too;
to show how much,
we repeat the whole procedure with 5 different random fold assignments.

```{r}
#| label: cv-mtcars
mtcars_degrees <- 1:4
n_repeats <- 5

set.seed(32)
mtcars_cv <-
  tidyr::expand_grid(repeat_id = seq_len(n_repeats), degree = mtcars_degrees) |>
  dplyr::group_by(repeat_id) |>
  dplyr::mutate(
    rmse = {
      folds <- assign_folds(nrow(mtcars), k = 4)
      purrr::map_dbl(
        degree,
        \(d) sqrt(cv_mse(mpg ~ poly(wt, d), mtcars, folds))
      )
    }
  ) |>
  dplyr::ungroup() |>
  tidyr::pivot_wider(
    names_from = degree,
    values_from = rmse,
    names_prefix = "degree_"
  )

mtcars_cv
```

:::{#tbl-cv-mtcars}

```{r}
#| label: cv-mtcars-best
mtcars_cv_best <-
  mtcars_cv |>
  tidyr::pivot_longer(
    -repeat_id,
    names_to = "degree",
    names_prefix = "degree_",
    values_to = "rmse"
  ) |>
  dplyr::group_by(repeat_id) |>
  dplyr::slice_min(rmse, n = 1) |>
  dplyr::ungroup()

mtcars_cv_best
```

Degree with the smallest 4-fold cross-validated RMSE
in each of `r n_repeats` random fold assignments
:::

```{r}
#| label: cv-mtcars-summary
mtcars_cv_range <-
  mtcars_cv |>
  dplyr::summarise(
    dplyr::across(dplyr::starts_with("degree_"), \(v) diff(range(v)))
  )
mtcars_chosen <- sort(unique(mtcars_cv_best$degree))
mtcars_choice_text <- if (length(mtcars_chosen) == 1) {
  paste0(
    "but every assignment chose degree ", mtcars_chosen,
    ": the estimates moved, and the winner did not"
  )
} else {
  paste0(
    "and the chosen degree was ",
    knitr::combine_words(mtcars_chosen, and = " or "),
    ", depending on how the cars happened to be split"
  )
}
mtcars_cv_range
```

Each row of the first table holds, for each degree,
the square root of $\cvest{4}$ (@eq-cv-estimate) for one fold assignment.
Across fold assignments,
the estimate for one degree varied by as much as
`r round(max(unlist(mtcars_cv_range)), 2)` mpg,
`r mtcars_choice_text`.
A single run of $k$-fold cross-validation
reports one of these rows,
so a small difference between two models' estimates
may not survive a different split.

:::
