<!--
This fragment's R chunks use objects made by earlier fragments
of model-validation.qmd, so it renders only after them:
rmse(), from _sol-overfitting.qmd.
-->
<!--
The section heading on ISLR p. 213 was checked in the PDF.
The four near-identical figure chunks of the rme version
are replaced by one plotting helper.
-->
:::{#exm-train-validation-test-split}
#### Numerical example

This example (@exm-train-validation-test-split)
uses R's built-in `mtcars` dataset
($n=32$ cars)
to predict fuel efficiency (`mpg`)
from vehicle weight (`wt`).
It uses one random split
to compare linear,
quadratic,
cubic,
and quartic models
(`mpg ~ wt`,
`mpg ~ wt + I(wt^2)`,
`mpg ~ wt + I(wt^2) + I(wt^3)`,
and `mpg ~ wt + I(wt^2) + I(wt^3) + I(wt^4)`)
on a validation set,
then reports the chosen model's test RMSE
on untouched test data
[@james2021islr2e, p. 213].

:::{#tbl-train-validation-test-split-sizes}
```{r}
#| code-fold: false
set.seed(108)
n <- nrow(mtcars)

n_train <- floor(0.6 * n)
n_valid <- floor(0.2 * n)

idx_train <- sample.int(n, size = n_train)
idx_remaining <- setdiff(seq_len(n), idx_train)
idx_valid <- sample(idx_remaining, size = n_valid)
idx_test <- setdiff(idx_remaining, idx_valid)

split_sizes <- tibble::tibble(
  split = c("training", "validation", "test"),
  n = c(length(idx_train), length(idx_valid), length(idx_test))
)

split_sizes
```

Sizes of the training,
validation,
and test splits
:::

```{r}
train_dat <- mtcars[idx_train, ]
valid_dat <- mtcars[idx_valid, ]
test_dat <- mtcars[idx_test, ]

model_linear <- lm(mpg ~ wt, data = train_dat)
model_quadratic <- lm(mpg ~ wt + I(wt^2), data = train_dat)
model_cubic <- lm(mpg ~ wt + I(wt^2) + I(wt^3), data = train_dat)
model_quartic <- lm(
  mpg ~ wt + I(wt^2) + I(wt^3) + I(wt^4),
  data = train_dat
)

candidate_models <- list(
  linear = model_linear,
  quadratic = model_quadratic,
  cubic = model_cubic,
  quartic = model_quartic
)

validation_results <- tibble::tibble(
  model = names(candidate_models)
) |>
  dplyr::mutate(
    training_RMSE = purrr::map_dbl(
      model,
      \(model_name) rmse(candidate_models[[model_name]], train_dat)
    ),
    validation_RMSE = purrr::map_dbl(
      model,
      \(model_name) rmse(candidate_models[[model_name]], valid_dat)
    )
  )

model_labels <- c(
  linear = "linear model",
  quadratic = "quadratic model",
  cubic = "cubic model",
  quartic = "quartic model"
)

chosen_model_name <-
  validation_results |>
  dplyr::arrange(validation_RMSE) |>
  dplyr::slice(1) |>
  dplyr::pull(model)

chosen_model_label <- model_labels[[chosen_model_name]]

chosen_model <- candidate_models[[chosen_model_name]]
chosen_model_validation_rmse <- rmse(chosen_model, valid_dat)

performance_comparison <- tibble::tibble(
  split = c("validation", "test"),
  model = chosen_model_name,
  RMSE = c(
    chosen_model_validation_rmse,
    rmse(chosen_model, test_dat)
  )
)

chosen_model_name
```

```{r}
#| label: setup-model-fit-figure-partitions
partition_colors <- c(
  training = "#1b9e77",
  validation = "#d95f02",
  test = "#7570b3"
)
partition_shapes <- c(
  training = 16,
  validation = 17,
  test = 15
)
wt_range <- range(c(train_dat$wt, valid_dat$wt, test_dat$wt))
mpg_range <- range(c(train_dat$mpg, valid_dat$mpg, test_dat$mpg))
wt_grid <- seq(wt_range[1], wt_range[2], length.out = 200)

partition_data <- dplyr::bind_rows(
  training = train_dat,
  validation = valid_dat,
  test = test_dat,
  .id = "series"
)

plot_partitions_with_fit <- function(model, fit_label) {
  series_levels <- c(names(partition_colors), fit_label)
  partition_data <- partition_data |>
    dplyr::mutate(series = factor(series, levels = series_levels))
  grid_data <- tibble::tibble(wt = wt_grid)
  fitted_curve <- tibble::tibble(
    wt = wt_grid,
    mpg = predict(model, newdata = grid_data),
    series = factor(fit_label, levels = series_levels)
  )
  # nolint start: object_usage_linter.
  ggplot2::ggplot(
    partition_data,
    ggplot2::aes(
      x = wt,
      y = mpg,
      color = series,
      shape = series
    )
  ) +
    ggplot2::geom_point() +
    ggplot2::geom_line(
      ggplot2::aes(linetype = series),
      data = fitted_curve,
      linewidth = 1
    ) +
    ggplot2::scale_color_manual(
      limits = series_levels,
      values = c(partition_colors, stats::setNames("blue", fit_label))
    ) +
    ggplot2::scale_shape_manual(
      limits = series_levels,
      values = c(partition_shapes, stats::setNames(NA, fit_label))
    ) +
    ggplot2::scale_linetype_manual(
      limits = series_levels,
      values = stats::setNames(c(0, 0, 0, 1), series_levels)
    ) +
    ggplot2::coord_cartesian(xlim = wt_range, ylim = mpg_range) +
    ggplot2::labs(color = NULL, shape = NULL, linetype = NULL) +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      legend.position = "inside",
      legend.position.inside = c(0.85, 0.8)
    )
  # nolint end
}

rbind(wt = wt_range, mpg = mpg_range)
```

:::{#fig-linear-model-superimposed}

```{r}
#| fig-alt: Scatterplot of miles per gallon versus vehicle weight.
#|   Points are colored and shaped by training,
#|   validation,
#|   and test partitions, with a legend naming them and the fit.
#|   The fitted linear regression line is superimposed.
#| code-fold: true
plot_partitions_with_fit(model_linear, "linear fit")
```

Linear model fit superimposed on data partitions
:::

:::{#fig-quadratic-model-superimposed}

```{r}
#| fig-alt: Scatterplot of miles per gallon versus vehicle weight.
#|   Points are colored and shaped by training,
#|   validation,
#|   and test partitions, with a legend naming them and the fit.
#|   The fitted quadratic curve is superimposed.
#| code-fold: true
plot_partitions_with_fit(model_quadratic, "quadratic fit")
```

Quadratic model fit superimposed on data partitions
:::

:::{#fig-cubic-model-superimposed}

```{r}
#| fig-alt: Scatterplot of miles per gallon versus vehicle weight.
#|   Points are colored and shaped by training,
#|   validation,
#|   and test partitions, with a legend naming them and the fit.
#|   The fitted cubic curve is superimposed.
#| code-fold: true
plot_partitions_with_fit(model_cubic, "cubic fit")
```

Cubic model fit superimposed on data partitions
:::

:::{#fig-quartic-model-superimposed}

```{r}
#| fig-alt: Scatterplot of miles per gallon versus vehicle weight.
#|   Points are colored and shaped by training,
#|   validation,
#|   and test partitions, with a legend naming them and the fit.
#|   The fitted quartic curve is superimposed.
#| code-fold: true
plot_partitions_with_fit(model_quartic, "quartic fit")
```

Quartic model fit superimposed on data partitions
:::

:::{#tbl-train-validation-results}
```{r}
validation_results
```

Training and validation RMSE for candidate models
:::

The chosen model is
the `r chosen_model_label`,
because it has the lowest validation RMSE of the four.
This split uses different cars from @exr-overfitting,
and its validation set has only 6 cars,
so the model it picks can differ from the one chosen there.

:::{#tbl-train-validation-test-rmse}
```{r}
performance_comparison
```

Validation and test RMSE
for the chosen model
:::

::: {.callout-note collapse="true"}
#### Source

The validation-set lab of @james2021islr2e [p. 213].
:::

:::
