:::{#sol-overfitting}

```{r}
#| label: overfitting-table
#| code-fold: false
train_cars <- mtcars[1:20, ]
held_out_cars <- mtcars[21:32, ]

prediction_mse <- function(model, data) {
  y <- data[[all.vars(stats::formula(model))[[1]]]]
  mean((predict(model, newdata = data) - y)^2)
}

rmse <- function(model, data) {
  sqrt(prediction_mse(model, data))
}

overfitting_results <- tibble::tibble(degree = 1:4) |>
  dplyr::mutate(
    fit = purrr::map(
      degree,
      \(d) lm(mpg ~ poly(wt, d, raw = TRUE), data = train_cars)
    ),
    training_RMSE = purrr::map_dbl(fit, rmse, data = train_cars),
    held_out_RMSE = purrr::map_dbl(fit, rmse, data = held_out_cars)
  ) |>
  dplyr::select(-fit)

overfitting_results
```

```{r}
#| label: overfitting-summary
best_train <- overfitting_results$degree[
  which.min(overfitting_results$training_RMSE)
]
best_held_out <- overfitting_results$degree[
  which.min(overfitting_results$held_out_RMSE)
]

c(best_train = best_train, best_held_out = best_held_out)
```

1. The smallest training RMSE is at degree `r best_train`.
   Each model contains the previous one as a special case,
   so its training RMSE cannot increase as the degree grows.

2. The smallest held-out RMSE is at degree `r best_held_out`.

3. No.
   Going from degree 2 to degree 3 or 4 lowers the training RMSE a little
   but leaves the held-out RMSE essentially unchanged
   (it does not fall below its degree-2 value).
   The terms beyond degree 2 improve the fit to the 20 training cars
   without improving predictions for the 12 held-out cars.
   Degree 1 to degree 2 is different:
   there the held-out RMSE also falls.

:::
